### 2023-06-06 submission_2/submission_2.qmd
Adds aurto files for submission 1 & 2

---
title: "Field training 02 - Leaf traits measurements"
subtitle: "Documentation"
subsubtitle: "Leaf traits measurements"
author: "Denis Streitmatter"
date: last-modified
format:
    pdf:
        documentclass: scrreport
        theme: united
        toc: true
        toc-depth: 2
        number-sections: true
        number-depth: 1
        toc-title: Table of Contents
        toc-location: body
        colorlinks: true
header-includes:  
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{{< pagebreak >}}

# Task 1: Leaf sampling

## Procedure

Date: 16.05.2023

Group-ID: 4

Participants: Denis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

Coordinates: 51.332222, 12.384167

Material used:

- Smartphone (GPS + plant species determination)
- Book (Flora Vegetativa)
- Survey sheet

{{< pagebreak >}}

## Leaf images

{{< pagebreak >}}

## Recorded species tables

{{< pagebreak >}}

<!--
Table in landscape format (uncomment if needed)

# landscape table format (better?)

::::::::{.landscape data-latex=""}

\begin{table}[]
\fontsize{22}{\baselineskip}\selectfont
\renewcommand{\arraystretch}{4}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\textbf{Group-ID} &
  \textbf{Plot-ID} &
  \textbf{Subplot-ID} &
  \textbf{Date} &
  \textbf{\begin{tabular}[c]{@{}c@{}}X-Coordinate \\ center\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Y-Coordinate \\ center\end{tabular}} &
  \textbf{Litter / \%} &
  \textbf{Soil / \%} &
  \textbf{Water / \%} &
  \textbf{Moss / \%} &
  \textbf{Height / cm} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Acer \\ platanoides\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Achillea \\ millefolium\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Bellis \\ perennis\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Capsella \\ bursa-pastoris\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Dactylis \\ glomerata\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Elymus \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ molle\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ pusillum\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Lolium \\ perenne\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Poa \\ annua\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Polygonum \\ aviculare\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Trifolium \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Veronica \\ arvensis\end{tabular}} \\ \hline
4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & 1.5 & 11 & 0 & 8 & 6 &     &     & 15 & 4 &   & 2  & 15 & 1 & 50 & 10 & 0.5 & 10 & 1   \\ \hline
4 & 1 & 2 & 16.05.2023 &           &           & 1   & 4  & 0 & 1 & 6 &     & 0.5 & 22 &   & 2 &    & 18 & 1 & 55 & 15 & 0.5 &    & 1   \\ \hline
4 & 1 & 3 & 16.05.2023 &           &           & 1   & 3  & 0 & 2 & 6 & 0.5 & 3   & 8  &   &   & 1  & 19 &   & 38 & 19 &     & 18 & 0.5 \\ \hline
4 & 1 & 4 & 16.05.2023 &           &           & 1   & 10 & 0 & 1 & 4 &     &     & 6  &   &   & 15 & 17 &   & 28 & 10 & 0.5 & 1  & 2   \\ \hline
\end{tabular}%
}
\caption{Subplot x species table for the recorded herbaceous vegetation. The values are the cover fraction of the total plot in {[}\%{]}.}
\end{table}
::::::::

-->

# Task 4: Leaf area measurement

There were two different approaches for measuring the leaf area: Using a specific tool or calculating it on the base of photos of the leafs on standardised A4 paper sheets

## Photo-based approach

For the photo-based approach we preprocessed the photos with the image-software `gimp` and then calculated the leaf size with a [python script](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/leaf_area_measurement.py).

### Preprocessing of the Photos

The goal of this step is to reduce the image to the necessary parts, mainly cutting out the A4 paper sheet. For this a few steps were used:

1. Load the image in gimp
2. Use the `lasso-select tool` for selecting the corners of the A4 paper
3. Press `CTRL + i` for inverting the selection
4. Press `CTRL + x` for cutting out everything thats not on the paper
5. Use `Image > Crop to content` to reduce the image size
    - this step is optional since alpha channel (= transparency) is ignored in the script, but this reduces file size and therefor makes it a bit faster
6. Save images in proper format for script:
    - Name of the file `{leaf-number}_someuniquename_{double}.jpg` (only use `double` if two A4 papers were used)
    - save all files of one species in one directory

### Calculating the leaf size

The code and the description how to use it can be found in [this git repository](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/). 
Here we will only describe the main ideas of the process. 
In general we know the area

First of all we hardcoded some values: The **A4 area** for reference,

# Woody vegetation sampling

## Procedure

Date: 16.05.2023

Group-ID: 4

Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

Coordinates: 51.332222, 12.384167

Material used:

- Quadrant
- Tape (for distance and diameter measurements)
- Smartphone (GPS + tree species determination)
- Survey sheet

Woody vegetation was surveyed using the 'point centered quarter method' (PCQM) on May 16th, 2023. This method is used to estimate plant abundance by placing a cross at a random location and assessing tree traits of the nearest tree in each quadrant. The data also serve as ground truthing data for 5x5 meter satellite images and are intended to represent species composition of the tree layer in described park area. 
For this purpose we selected again a random point by throwing a pile in the air and noted down the metadata, i.e. date and coordinates in LatLong/WGS84 format using a smartphone. No explicit photos of the recorded location nor trees were taken.
In order to record the tree layer using the PCQM, we identified the nearest tree in each quadrant using a tape for the distance measure and knowledge as well as the software 'FloraIncognita' to determine the species. The nearest species were named Acer platanoides, Acer pseudoplatanus, Cornus sanguinea and Quercus robur. For each tree, the distance from the determined center location and the girth at breast height (~130cm) [cm] were measured with a tape. The diameter at breast height [cm] was then calculated with $\frac{girth [cm]}{\pi}$. Based on the diameter, the basal area [$\text{cm}^2$] was calculated with the formula $\pi \text{r}^2$, with $r = \frac{diameter [cm]}{2} [cm]$.
Afterwards, the recorded data as well as the metadata were entered in a GoogleDocs sheet, which can be see in Table 2.1.

{{< pagebreak >}}

## Recorded species table

\begin{table}[h!]
\fontsize{32}{\baselineskip}\selectfont
\renewcommand{\arraystretch}{8}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
\hline
\textbf{Group-ID} &
  \textbf{Plot-ID} &
  \textbf{Tree-ID} &
  \textbf{Date} &
  \textbf{\begin{tabular}[c]{@{}l@{}}X-Coordinate\\ center\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}l@{}}Y-Coordinate\\ center\end{tabular}} &
  \textbf{Species} &
  \textbf{Distance} &
  \textbf{Girth (cm)} &
  \textbf{DBH / cm} &
  \textbf{BA / cm²} &
  \textbf{total BA / cm²} \\ \hline
4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & Acer platanoides    & 9.8  & 239 & 76.08 & 4545.54 & 4545.54 \\ \hline
4 & 1 & 2 & 16.05.2023 &           &           & Acer pseudoplatanus & 10.9 & 205 & 65.25 & 3344.24 & 3344.24 \\ \hline
4 & 1 & 3 & 16.05.2023 &           &           & Cornus sanguinea    & 15.7 & 36  & 11.46 & 103.13  & 103.13  \\ \hline
4 & 1 & 4 & 16.05.2023 &           &           & Quercus robur       & 25.2 & 302 & 96.13 & 7257.78 & 7257.78 \\ \hline
\end{tabular}%
}
\caption{Table for the recorded woody vegetation including metadata.}
\end{table}

### 2023-06-06 submission_2/submission_2.qmd
updated task4

@@
 
+In the next step, we determined the mass per leaf area.
+We took the sheets out of the paper bag and placed them one by one into the sheet scanner.
+We had to make sure that the sheets were fully spread out, did not overlap, and were within the measurement range area.
+For some sheets, the scanner was not large enough to measure all 10 sheets at once.
+So we did multiple scans and added up and documented the measured leaf area for each species.
+
+To check the correctness of the scanner we performed a second type of leaf surface measurement.
+For this, we placed the leaves on a DinA4 sheet and took a photo.
+We wrote a small script to convert the scans of sheets and paper into black and white images and calculate the cumulative sheet area from the proportion of black pixels of the total paper area.
+Then we compared the measured area with the result of the sheet scanner.
+They were very similar everywhere with slight variations (see documentation).
 There were two different approaches for measuring the leaf area: Using a specific tool or calculating it on the base of photos of the leafs on standardised A4 paper sheets
@@
 First of all we hardcoded some values: The **A4 area** for reference,
+
+## LMA
+
+To calculate the biomass of the sheets, we now had to determine the dry weight of the sheets.
+To do this, we put the leaves back in the paper bag, and put them in the drying oven at 50 - 55 °C.
+The samples were dried for 72h.
+Then we measured the dry weight and noted the values.
+Finally we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Aria (LMA).
 

### 2023-06-06 submission_2/submission_2.qmd
Adds task 4 script description

@@
 Here we will only describe the main ideas of the process. 
-In general we know the area
-
-First of all we hardcoded some values: The **A4 area** for reference,
+In general we know the area of the A4 paper, so we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` and therefor we can calculate the area of the leaf. 
+Unfortunately shadows of not fully flat leaves can add a lot of noise to the trivial approach (converting it to greyscale and cutting at a certain threshold). 
+To fix this we transposed the image in the `HSV` color space, where all grey/white colors have a very low saturation value (compared to the green leaves), so we determine a  saturation threshold here (a comparison feature between Bitmask, HSV and original image was implemented to find a fitting value) and generate the bitmask in this way.
+The transparent parts are not considered during the area calculation, only the white (paper), grey (shadows, counted as paper) and green (leaf) pixels.
 
@@
 Finally we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Aria (LMA).
-
-# Woody vegetation sampling
-
-## Procedure
-
-Date: 16.05.2023
-
-Group-ID: 4
-
-Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell
-
-Coordinates: 51.332222, 12.384167
-
-Material used:
-
-- Quadrant
-- Tape (for distance and diameter measurements)
-- Smartphone (GPS + tree species determination)
-- Survey sheet
-
-Woody vegetation was surveyed using the 'point centered quarter method' (PCQM) on May 16th, 2023.
-This method is used to estimate plant abundance by placing a cross at a random location and assessing tree traits of the nearest tree in each quadrant.
-The data also serve as ground truthing data for 5x5 meter satellite images and are intended to represent species composition of the tree layer in described park area.
-For this purpose we selected again a random point by throwing a pile in the air and noted down the metadata, i.e.
-date and coordinates in LatLong/WGS84 format using a smartphone.
-No explicit photos of the recorded location nor trees were taken.
-In order to record the tree layer using the PCQM, we identified the nearest tree in each quadrant using a tape for the distance measure and knowledge as well as the software 'FloraIncognita' to determine the species.
-The nearest species were named Acer platanoides, Acer pseudoplatanus, Cornus sanguinea and Quercus robur.
-For each tree, the distance from the determined center location and the girth at breast height (~130cm) [cm] were measured with a tape.
-The diameter at breast height [cm] was then calculated with $\frac{girth [cm]}{\pi}$.
-Based on the diameter, the basal area [$\text{cm}^2$] was calculated with the formula $\pi \text{r}^2$, with $r = \frac{diameter [cm]}{2} [cm]$.
-Afterwards, the recorded data as well as the metadata were entered in a GoogleDocs sheet, which can be see in Table 2.1.
-
-{{< pagebreak >}}
-
-## Recorded species table
-
-\begin{table}[h!]
-\fontsize{32}{\baselineskip}\selectfont
-\renewcommand{\arraystretch}{8}
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
-\hline
-\textbf{Group-ID} &
-  \textbf{Plot-ID} &
-  \textbf{Tree-ID} &
-  \textbf{Date} &
-  \textbf{\begin{tabular}[c]{@{}l@{}}X-Coordinate\\ center\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}l@{}}Y-Coordinate\\ center\end{tabular}} &
-  \textbf{Species} &
-  \textbf{Distance} &
-  \textbf{Girth (cm)} &
-  \textbf{DBH / cm} &
-  \textbf{BA / cm²} &
-  \textbf{total BA / cm²} \\ \hline
-4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & Acer platanoides    & 9.8  & 239 & 76.08 & 4545.54 & 4545.54 \\ \hline
-4 & 1 & 2 & 16.05.2023 &           &           & Acer pseudoplatanus & 10.9 & 205 & 65.25 & 3344.24 & 3344.24 \\ \hline
-4 & 1 & 3 & 16.05.2023 &           &           & Cornus sanguinea    & 15.7 & 36  & 11.46 & 103.13  & 103.13  \\ \hline
-4 & 1 & 4 & 16.05.2023 &           &           & Quercus robur       & 25.2 & 302 & 96.13 & 7257.78 & 7257.78 \\ \hline
-\end{tabular}%
-}
-\caption{Table for the recorded woody vegetation including metadata.}
-\end{table}

### 2023-06-06 leaf_area_measure/leaf_area_measurement.py
Small docu adjustment

@@ -57,7 +57,7 @@ def calculate_area_relation_of_mask(mask: np.ndarray) -> float:
 class LeafAreaMeasurer:
 
     def __init__(self, saturation_threshold: float = 0.2, skipcolors: List = [ALPHA_VALUE_HSV], max_skipping_distance: float = 0.1) -> None:
-        """Class for calculating the Area of images. 
+        """Class for calculating the Area of leaf images. 
         saturation_threshold: Is the threshold at which something is not considered a 'leaf' anymore
         skipcolors: List of hsv colors (as triple) which are not part of the area (e.g. the alpha channel)
         max_skipping_distance: Is the distance a pixel can be away from the skipping color to be considered skipping"""

### 2023-06-06 submission_2/submission_2.qmd
Some small improvements

@@
 
-For the photo-based approach we preprocessed the photos with the image-software `gimp` and then calculated the leaf size with a [python script](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/leaf_area_measurement.py).
+For this approach we took Photos of single or multiple leafs on one (or two) A4 papers. 
+These photos were then preprocessed with the image-software `gimp` and then calculated the leaf size with a [python script](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/leaf_area_measurement.py).
 
@@
 
-The goal of this step is to reduce the image to the necessary parts, mainly cutting out the A4 paper sheet. For this a few steps were used:
+The goal of this step is to reduce the image to the necessary parts, mainly cutting out the A4 paper sheet. For this a few steps were performed:
 
@@
 Here we will only describe the main ideas of the process. 
-In general we know the area of the A4 paper, so we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` and therefor we can calculate the area of the leaf. 
+In general we know the area of the A4 paper, so to determine leaf size we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` and therefor we can calculate the area of the leaf.
 Unfortunately shadows of not fully flat leaves can add a lot of noise to the trivial approach (converting it to greyscale and cutting at a certain threshold). 
-To fix this we transposed the image in the `HSV` color space, where all grey/white colors have a very low saturation value (compared to the green leaves), so we determine a  saturation threshold here (a comparison feature between Bitmask, HSV and original image was implemented to find a fitting value) and generate the bitmask in this way.
+To fix this we transposed the image in the `HSV` color space, where all grey/white colors have a very low saturation value (compared to the green leaves), so we determine a  saturation threshold here (a comparison feature between Bitmask, `HSV` and original image was implemented to find a fitting value) and generate the bitmask in this way.
 The transparent parts are not considered during the area calculation, only the white (paper), grey (shadows, counted as paper) and green (leaf) pixels.

### 2023-06-12 submission_1/submission_1.qmd
Adds submission 1 quarto project

---
title: "Field training 01 - Vegetation sampling"
subtitle: "Documentation"
subsubtitle: "Vegetation sampling"
author: "Denis Streitmatter"
date: last-modified
format:
    pdf:
        documentclass: scrreport
        theme: united
        toc: true
        toc-depth: 2
        number-sections: true
        number-depth: 1
        toc-title: Table of Contents
        toc-location: body
        colorlinks: true
header-includes:  
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{{< pagebreak >}}

# Herbaceous vegetation sampling

## Procedure

Date: 16.05.2023

Group-ID: 4

Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

Coordinates: 51.332222, 12.384167

Material used:

- Frame ($1 \, \text{m}^2$)
- Peg and 3 m cord
- Smartphone (GPS + plant species determination)
- Book (Flora Vegetativa)
- Survey sheet

The aim of the vegetation sampling was to collect a random vegetation ground truth sample for 5x5 meter remote sensing imagery. On May 16th 2023, we recorded herbaceous vegetation on four subplots of 1x1 meter each ($1 \, \text{m}^2$) on the meadows of a small park named 'Kanonenteichanlage' between Liebigstraße and Brüderstraße in 04103 Leipzig, Germany. 
Firstly, we selected a homogeneous location without trees, shrubs or disturbances within a range of 20 meters. On that area, the location where a randomly thrown pile landed marked our central subplot. Secondly, we measured the coordinates of the relevé in LatLong/WGS84 format using a smartphone. Following, we recorded metadata as well as plant species abundance for the present herbaceous vegetation. The metadata includes a plot image, coverage of litter [%], soil [%], water [%] and moss [%] as well as the height. In order to assess the height, we measured the vertical extend from the base of the plant (soil level) to the tip of the highest point using a ruler, and excluding any aerial leaves or roots. This process was repeated several times (~20) for different species and the mean height calculated subsequently. For the remaining metadata, we calculated the mean of our personal/individual estimates (N = 6). Lastly, we identified all present species using the free software 'FloraIncognita' and the book 'Flora Vegetativa - Ein Bestimmungsbuch für Pflanzen der Schweiz im blütenlosen Zustand' (2020) by Eggenberg and Möhl. The cover fraction of each species was again calculated by taking the mean of our six individual estimates.
The same procedure was conducted for three more subplots arranged in 120° directions three meters from the center subplot.
Additionally, the recorded species cover fractions as well as the metadata of the four subplots was aggregated to a main plot using the mean of the four values, respectively.
Afterwards, the recorded data as well as the metadata as well as the aggregated data at the plot level were entered in GoogleDocs sheets, which can be see in Tables 1.1 and 1.2.

{{< pagebreak >}}

## Plot images
![Subplot 1 (center plot). Photo taken at: 05/16/2023 13:32 CET.](plot_photos/plot1.jpeg "Subplot 1 (center plot). Photo taken at: 05/16/2023 13:32 CET.")

![Subplot 2. Photo taken at: 05/16/2023 14:37 CET.](plot_photos/plot2.jpeg "Subplot 2. Photo taken at: 05/16/2023 14:37 CET.")

![Subplot 3. Photo taken at: 05/16/2023 15:04 CET.](plot_photos/plot3.jpeg "Subplot 3. Photo taken at: 05/16/2023 15:04 CET.")

![Subplot 4. Photo taken at: 05/16/2023 15:14 CET.](plot_photos/plot4.jpeg "Subplot 4. Photo taken at: 05/16/2023 15:14 CET.")

{{< pagebreak >}}

## Recorded species tables

\begin{table}[h!]
\fontsize{22}{\baselineskip}\selectfont
\renewcommand{\arraystretch}{4}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\textbf{Group-ID} &
  \textbf{Plot-ID} &
  \textbf{Subplot-ID} &
  \textbf{Date} &
  \textbf{\begin{tabular}[c]{@{}c@{}}X-Coordinate \\ center\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Y-Coordinate \\ center\end{tabular}} &
  \textbf{Litter / \%} &
  \textbf{Soil / \%} &
  \textbf{Water / \%} &
  \textbf{Moss / \%} &
  \textbf{Height / cm} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Acer \\ platanoides\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Achillea \\ millefolium\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Bellis \\ perennis\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Capsella \\ bursa-pastoris\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Dactylis \\ glomerata\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Elymus \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ molle\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ pusillum\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Lolium \\ perenne\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Poa \\ annua\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Polygonum \\ aviculare\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Trifolium \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Veronica \\ arvensis\end{tabular}} \\ \hline
4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & 1.5 & 11 & 0 & 8 & 6 &     &     & 15 & 4 &   & 2  & 15 & 1 & 50 & 10 & 0.5 & 10 & 1   \\ \hline
4 & 1 & 2 & 16.05.2023 &           &           & 1   & 4  & 0 & 1 & 6 &     & 0.5 & 22 &   & 2 &    & 18 & 1 & 55 & 15 & 0.5 &    & 1   \\ \hline
4 & 1 & 3 & 16.05.2023 &           &           & 1   & 3  & 0 & 2 & 6 & 0.5 & 3   & 8  &   &   & 1  & 19 &   & 38 & 19 &     & 18 & 0.5 \\ \hline
4 & 1 & 4 & 16.05.2023 &           &           & 1   & 10 & 0 & 1 & 4 &     &     & 6  &   &   & 15 & 17 &   & 28 & 10 & 0.5 & 1  & 2   \\ \hline
\end{tabular}%
}
\caption{Subplot x species table for the recorded herbaceous vegetation including metadata. The values are the cover fraction of the total plot in {[}\%{]}.}
\end{table}

\begin{table}[h!]
\fontsize{32}{\baselineskip}\selectfont
\renewcommand{\arraystretch}{8}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\textbf{Group-ID} &
  \textbf{Plot-ID} &
  \textbf{Date} &
  \textbf{X-Coordinate} &
  \textbf{Y-Coordinate} &
  \textbf{Litter / \%} &
  \textbf{Soil / \%} &
  \textbf{Water / \%} &
  \textbf{Moss / \%} &
  \textbf{Height / cm} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Acer \\ platanoides\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Achillea \\ millefolium\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Bellis \\ perennis\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Capsella \\ bursa-pastoris\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Dactylis \\ glomerata\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Elymus \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ molle\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ pusillum\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Lolium \\ perenne\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Poa \\ annua\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Polygonum \\ aviculare\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Trifolium \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Veronica \\ arvensis\end{tabular}} \\ \hline
4 &
  1 &
  16.05.2023 &
  51.332222 &
  12.384167 &
  1.13 &
  7 &
  0 &
  3 &
  5.5 &
  0.13 &
  0.88 &
  12.75 &
  1 &
  0.5 &
  4.5 &
  17.25 &
  0.5 &
  42.75 &
  13.5 &
  0.38 &
  7.25 &
  4.13 \\ \hline
\end{tabular}%
}
\caption{Aggregated plot x species table for the recorded herbaceous vegetation including metadata. The values are the mean cover fraction of the four subplots plot in {[}\%{]}.}
\end{table}

{{< pagebreak >}}

<!--
Table in landscape format (uncomment if needed)

# landscape table format (better?)

::::::::{.landscape data-latex=""}

\begin{table}[]
\fontsize{22}{\baselineskip}\selectfont
\renewcommand{\arraystretch}{4}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\textbf{Group-ID} &
  \textbf{Plot-ID} &
  \textbf{Subplot-ID} &
  \textbf{Date} &
  \textbf{\begin{tabular}[c]{@{}c@{}}X-Coordinate \\ center\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Y-Coordinate \\ center\end{tabular}} &
  \textbf{Litter / \%} &
  \textbf{Soil / \%} &
  \textbf{Water / \%} &
  \textbf{Moss / \%} &
  \textbf{Height / cm} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Acer \\ platanoides\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Achillea \\ millefolium\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Bellis \\ perennis\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Capsella \\ bursa-pastoris\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Dactylis \\ glomerata\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Elymus \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ molle\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ pusillum\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Lolium \\ perenne\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Poa \\ annua\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Polygonum \\ aviculare\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Trifolium \\ repens\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}c@{}}Veronica \\ arvensis\end{tabular}} \\ \hline
4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & 1.5 & 11 & 0 & 8 & 6 &     &     & 15 & 4 &   & 2  & 15 & 1 & 50 & 10 & 0.5 & 10 & 1   \\ \hline
4 & 1 & 2 & 16.05.2023 &           &           & 1   & 4  & 0 & 1 & 6 &     & 0.5 & 22 &   & 2 &    & 18 & 1 & 55 & 15 & 0.5 &    & 1   \\ \hline
4 & 1 & 3 & 16.05.2023 &           &           & 1   & 3  & 0 & 2 & 6 & 0.5 & 3   & 8  &   &   & 1  & 19 &   & 38 & 19 &     & 18 & 0.5 \\ \hline
4 & 1 & 4 & 16.05.2023 &           &           & 1   & 10 & 0 & 1 & 4 &     &     & 6  &   &   & 15 & 17 &   & 28 & 10 & 0.5 & 1  & 2   \\ \hline
\end{tabular}%
}
\caption{Subplot x species table for the recorded herbaceous vegetation. The values are the cover fraction of the total plot in {[}\%{]}.}
\end{table}
::::::::

-->

# Woody vegetation sampling

## Procedure

Date: 16.05.2023

Group-ID: 4

Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

Coordinates: 51.332222, 12.384167

Material used:

- Quadrant
- Tape (for distance and diameter measurements)
- Smartphone (GPS + tree species determination)
- Survey sheet

Woody vegetation was surveyed using the 'point centered quarter method' (PCQM) on May 16th, 2023. This method is used to estimate plant abundance by placing a cross at a random location and assessing tree traits of the nearest tree in each quadrant. The data also serve as ground truthing data for 5x5 meter satellite images and are intended to represent species composition of the tree layer in described park area. 
For this purpose we selected again a random point by throwing a pile in the air and noted down the metadata, i.e. date and coordinates in LatLong/WGS84 format using a smartphone. No explicit photos of the recorded location nor trees were taken.
In order to record the tree layer using the PCQM, we identified the nearest tree in each quadrant using a tape for the distance measure and knowledge as well as the software 'FloraIncognita' to determine the species. The nearest species were named Acer platanoides, Acer pseudoplatanus, Cornus sanguinea and Quercus robur. For each tree, the distance from the determined center location and the girth at breast height (~130cm) [cm] were measured with a tape. The diameter at breast height [cm] was then calculated with $\frac{girth [cm]}{\pi}$. Based on the diameter, the basal area [$\text{cm}^2$] was calculated with the formula $\pi \text{r}^2$, with $r = \frac{diameter [cm]}{2} [cm]$.
Afterwards, the recorded data as well as the metadata were entered in a GoogleDocs sheet, which can be see in Table 2.1.

{{< pagebreak >}}

## Recorded species table

\begin{table}[h!]
\fontsize{32}{\baselineskip}\selectfont
\renewcommand{\arraystretch}{8}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
\hline
\textbf{Group-ID} &
  \textbf{Plot-ID} &
  \textbf{Tree-ID} &
  \textbf{Date} &
  \textbf{\begin{tabular}[c]{@{}l@{}}X-Coordinate\\ center\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}l@{}}Y-Coordinate\\ center\end{tabular}} &
  \textbf{Species} &
  \textbf{Distance} &
  \textbf{Girth (cm)} &
  \textbf{DBH / cm} &
  \textbf{BA / cm²} &
  \textbf{total BA / cm²} \\ \hline
4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & Acer platanoides    & 9.8  & 239 & 76.08 & 4545.54 & 4545.54 \\ \hline
4 & 1 & 2 & 16.05.2023 &           &           & Acer pseudoplatanus & 10.9 & 205 & 65.25 & 3344.24 & 3344.24 \\ \hline
4 & 1 & 3 & 16.05.2023 &           &           & Cornus sanguinea    & 15.7 & 36  & 11.46 & 103.13  & 103.13  \\ \hline
4 & 1 & 4 & 16.05.2023 &           &           & Quercus robur       & 25.2 & 302 & 96.13 & 7257.78 & 7257.78 \\ \hline
\end{tabular}%
}
\caption{Table for the recorded woody vegetation including metadata.}
\end{table}

### 2023-06-12 submission_1/submission_1.qmd
restructures submissions
Also some tiny changes to the submissions

@@
 For the remaining metadata, we calculated the mean of our personal/individual estimates (N = 6).
-Lastly, we identified all present species using the free software 'FloraIncognita' and the book 'Flora Vegetativa - Ein Bestimmungsbuch für Pflanzen der Schweiz im blütenlosen Zustand' (2020) by Eggenberg and Möhl.
+Lastly, we identified all present species using the free software 'FloraIncognita', 'PlantNet' and the book 'Flora Vegetativa - Ein Bestimmungsbuch für Pflanzen der Schweiz im blütenlosen Zustand' (2020) by Eggenberg and Möhl.
 The cover fraction of each species was again calculated by taking the mean of our six individual estimates.

### 2023-06-12 submission_2/submission_2.qmd
restructures submissions
Also some tiny changes to the submissions

@@
 In the next step, we determined the mass per leaf area.
+For this task two different methods were used: A specialized scanner for this exact purpose and a python script calculating the leaf area based of photos on an a DIN-A4 paper.
+
+## Specialized Scanner
+
  We took the sheets out of the paper bag and placed them one by one into the sheet scanner.
@@
 These photos were then preprocessed with the image-software `gimp` and then calculated the leaf size with a [python script](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/leaf_area_measurement.py).
+The photos (unprocessed and otherwise) can be found [here](https://github.com/yum-yab/ground_truthing_module/tree/main/leaf_area_measure/leaf_images)
 
@@
 Here we will only describe the main ideas of the process. 
-In general we know the area of the A4 paper, so to determine leaf size we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` and therefor we can calculate the area of the leaf.
+Since we know the area of the A4 paper, to determine leaf size we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` ($\equiv$ DIN-A4) and therefor we can calculate the area of the leaf.
 Unfortunately shadows of not fully flat leaves can add a lot of noise to the trivial approach (converting it to greyscale and cutting at a certain threshold). 

### 2023-06-13 submission_2/submission_2.qmd

@@
 - Survey sheet
+
+{{< pagebreak >}}
+
+# Task 2: Leaf chlorophyll measurements
+## Procedure
+
+Material used:
+
+- Leaf samples of 8 plant species
+- Konica Minolta SPAD 502
+- Leaf trait sheet of paper
+
+The aim of this measurement was to determine the leaf chlorophyll content (per area) using the *Konica Minolta SPAD-502 Plus* device.
+This instrument measures leaf absorbance in red and near-infrared regions using two LEDs at peak wavelengths of 650 nm and 940 nm.
+It calculates a Soil Plant Analysis Development (SPAD) value by dividing the light transmission intensities and determines the relative chlorophyll content in the sample leaf (Konica Minolta Optics, 2012).
+A more detailed description of the instrument can be found under [Süß, A., Danner, M., Obster, C., Locherer, M., Hank, T., Richter, K.
+(2015): Measuring Leaf Chlorophyll Content with the Konica Minolta SPAD-502Plus – Theory, Measurement, Problems, Interpretation.
+EnMAP Field Guides Technical Report, GFZ Data Services](https://gfzpublic.gfz-potsdam.de/rest/items/item_1388302/component/file_1388303/content) (last accessed: 06/06/2023).
+
+We measured the chlorophyll content on the upper side of each leaf in four (*Geranium molle, G.
+sanguineum*) to six spots (*Acer platanoides, Alliaria petiolata, Ballota nigra, Geum urbanum, Philadelphus Coronarius, Spirae Chamaedryfolia*), depending on the leaf size and shape.
+During the measurement, we padi attention so that no obvious outliers were recorded, no veins were measured and that the optical window of the instrument was fully covered by the leaf.
+For each of the ten sampled leaves, the mean of the four to six measurements was written down on the leaf trait sheet (see columns *SPAD1...SPAD10* of table < INSERT TABLE NUMBERING>).
+Afterwards, the mean and standard deviation for all the leaves of each species were calculated.
+
+## Chlorophyll conversions
+
+In order to convert the mean values per species measured with the SPAD-instrument, the following equations from *Süß et al. (2015)* were used:
+
+![Table with formulas to convert measured SPAD pigment values to chlorophyll per area values.](SPAD_conversion_table.png "Table with formulas to convert measured SPAD pigment values to chlorophyll per area values.")
+
+The results of the conversions can be seen in the attached data table as well as in the figure below.
+
+![Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.](SPAD_conversion_variation.png "Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.")
 

### 2023-06-13 submission_2/submission_2.qmd
Intro, Task 3,5 and tables

@@
 
-# Task 1: Leaf sampling
+# Leaf Properties Sampling
 
-## Procedure
+# Task 1: Procedure
 
-Date: 16.05.2023
+<!-- Task1: Leaf sampling in the park and quick run through of all processing steps -->
+
+Date: 23.05.2023
 
@@
 
-Participants: Denis, Fabian, Louis, Lukas, Marie-Louise, Maxwell
+Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell
 
 Coordinates: 51.332222, 12.384167
+
+Material used for Sampling:
+
+- Paper Bags and Pencil
+- Smartphone (GPS + plant species determination)
+- Book (Flora Vegetativa)
+- Survey sheet
+
+Materials used for Analysis:
+
+- Soil Plant Analysis Development (SPAD) meter for chlorophyll measurements
+- Precision balance
+- Leaf scanner
+- A4 paper and camera
+- Drying cabinet
+
+The aim of the leaf property sampling was to collect and analyse leaf traits of different plant species.
+On May 23rd 2023, we collected leaf samples of eight different species in Kanonenteichanlage, a small park between Liebigstraße and Brüderstraße in 04103 Leipzig, Germany.
+
+The plant species that were sampled were choosen based on their abundance in the park.
+For each species 10 leaves from different individuals were sampled.
+When sampling leaves we were careful to sample the whole leaf.
+Leaves were chosen based on the following criteria, healthy, fully grown and from a sun-lit environment where possible.
+We sampled trees, shrubs and herbs.
+The plants were identified using the smartphone app 'FloraIncognita' and the latin names of the species were recorded in a table.
+The leaves were collected in labeled paper bags and taken back to the seminar room for further analyisis.
+The leaf traits recorded for each species were fresh weight, dry weight, leaf area, and chlorophyll content.
+The fresh weight was determined first using a precision balance.
+In the next step chlorophyll concentration was derived from plant greenness using a Soil Plant Analysis Development (SPAD) meter.
+We used the leaf scanner to measure leaf area, we also took pictures of the leafs on an A4 sheet to derive leaf area using a script for image analysis.
+The leaves were then put in a drying cabinate at 50° to 55° for 48h.
+In a last step dry weight was determined.
+All the data was recorded in a google sheet.
+
+{{< pagebreak >}}
+
+## Leaf water content
+
+<!-- Task 3 and 5 -->
+The total fresh weight and dry weight of the leave samples were used to determine the equivalent water thickness (EWT, @tbl-data-summary).
+The total weight of the leaves was determined through carefully placing the samples on precision laboratory balances.
+The equivalent water thickness is calculated by substracting the weight of the dry sample from the weight of the fresh sample and then deviding by the product of the density of water and the leaf area (@eq-EWT).
+
+ $$
+ EWT = \frac{fresh\: weight\: [mg]-dry\: weight\: [mg]}{1000\:\frac{mg}{cm^{3}}\; \; leaf\: area\: [cm^{2}]}
+ $$ {#eq-EWT}
+
+{{< pagebreak >}}
+
+<!-- change to landscape -->
+\KOMAoptions{paper=landscape,pagesize} 
+\recalctypearea
+
+<!-- please feel free to make these tables look prettier-->
+
+| Species | Leaf area (scanner) cm² | Leaf area (camera) cm² | Fresh weight [mg] | Dry weight [mg] | LMA [mg/cm²] scanner | LMA [mg/cm²] camera | EWT [mg/cm²] scanner | EWT [mg/cm²] camera |
+|---|---|---|---|---|---|---|---|---|
+| Philadelphus Coronarius | 719.97 | 751.19 | 13830 | 2440 | 19.20913371 | 18.41078822 | 0.02 | 0.02 |
+| Spirae Chamaedryfolia | 128.27 | 122.26 | 1760 | 620 | 13.72105715 | 14.39555047 | 0.01 | 0.01 |
+| Geranium molle | 119.11 | 141.13 | 3590 | 590 | 30.14020653 | 25.43753986 | 0.03 | 0.02 |
+| Alliaria petiolata | 810.5 | 892.12 | 17720 | 3630 | 21.8630475 | 19.86279873 | 0.02 | 0.02 |
+| Geum urbanum | 184.8 | 194.62 | 2390 | 860 | 12.93290043 | 12.28034118 | 0.01 | 0.01 |
+| Ballota nigra | 105.88 | 110.4 | 2090 | 610 | 19.73932754 | 18.93115942 | 0.01 | 0.01 |
+| Geranium sanguineum | 82.62 | 97.56 | 1910 | 530 | 23.11788913 | 19.57769578 | 0.02 | 0.01 |
+| Acer platanoides | 1142.2 | 1343.79 | 15110 | 4530 | 13.22885659 | 11.24431645 | 0.01 | 0.01 |
+
+: Leaf properties: leaf area, leaf fresh and dry weight, LMA (leaf mass per area) and EWT (equivalent water thickness) {#tbl-data-summary}
+
+{{< pagebreak >}}
+
+| Species | SPAD mean | SPAD stdv | chl Eq. 1 | chl Eq. 2 | chl Eq. 3 | chl Eq. 4 | chl Eq. 5 |
+|---|---|---|---|---|---|---|---|
+| Philadelphus Coronarius | 47.65 | 1.681434573 | 265.467675 | 621.573065 | 608.2915624 | 61.56568402 | 62.75488428 |
+| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 243.8084 | 584.63104 | 564.7495158 | 56.03685716 | 56.72252579 |
+| Geranium molle | 36.1 | 3.919183588 | 155.5059 | 425.94494 | 386.0649413 | 36.23261347 | 35.51025089 |
+| Alliaria petiolata | 27.64 | 1.416725489 | 93.583776 | 301.9521344 | 257.008503 | 24.57293509 | 23.40017292 |
+| Geum urbanum | 28.03 | 2.746330886 | 96.092139 | 307.3093226 | 262.3688604 | 25.01677595 | 23.85444116 |
+| Ballota nigra | 38.44 | 4.576558629 | 175.413216 | 463.1218304 | 426.6499938 | 40.34089174 | 39.8524312 |
+| Geranium sanguineum | 48.9 | 6.534332407 | 279.1283 | 644.56894 | 635.7441646 | 65.20131502 | 66.7438002 |
+| Acer platanoides | 31.59 | 5.872998477 | 120.535811 | 357.8139034 | 313.8933755 | 29.4575616 | 28.43095204 |
+
+: Leaf properties: SPAD measurements and chlorophyll concentration from 5 different equations {#tbl-data-summary}
+
+{{< pagebreak >}}
+
+<!-- change back to portrait -->
+\KOMAoptions{paper=portrait,pagesize} 
+\recalctypearea
 

### 2023-06-13 submission_2/submission_2.qmd
Adds authors, fixes order and naming

@@
 subsubtitle: "Leaf traits measurements"
-author: "Denis Streitmatter"
+author: 
+    - "Group 4" 
+    - "Denis Streitmatter"
+    - "Marie-Louise Korte"
+    - "Louis Trinkle"
+    - "Maxwell Sivertsen"
+    - "Fabian Götz"
+    - "Lukas Erzfeld"
 date: last-modified
@@
         toc-depth: 2
-        number-sections: true
+        number-sections: false 
         number-depth: 1
@@
 
-# Leaf Properties Sampling
-
-# Task 1: Procedure
+# Task 1: Leaf sampling
 
@@
 
-Group-ID: 4
-
-Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell
+<!-- Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell -->
 
@@
 
-## Leaf water content
-
-<!-- Task 3 and 5 -->
-The total fresh weight and dry weight of the leave samples were used to determine the equivalent water thickness (EWT, @tbl-data-summary).
-The total weight of the leaves was determined through carefully placing the samples on precision laboratory balances.
-The equivalent water thickness is calculated by substracting the weight of the dry sample from the weight of the fresh sample and then deviding by the product of the density of water and the leaf area (@eq-EWT).
-
- $$
- EWT = \frac{fresh\: weight\: [mg]-dry\: weight\: [mg]}{1000\:\frac{mg}{cm^{3}}\; \; leaf\: area\: [cm^{2}]}
- $$ {#eq-EWT}
-
-{{< pagebreak >}}
-
-<!-- change to landscape -->
-\KOMAoptions{paper=landscape,pagesize} 
-\recalctypearea
-
-<!-- please feel free to make these tables look prettier-->
-
-| Species | Leaf area (scanner) cm² | Leaf area (camera) cm² | Fresh weight [mg] | Dry weight [mg] | LMA [mg/cm²] scanner | LMA [mg/cm²] camera | EWT [mg/cm²] scanner | EWT [mg/cm²] camera |
-|---|---|---|---|---|---|---|---|---|
-| Philadelphus Coronarius | 719.97 | 751.19 | 13830 | 2440 | 19.20913371 | 18.41078822 | 0.02 | 0.02 |
-| Spirae Chamaedryfolia | 128.27 | 122.26 | 1760 | 620 | 13.72105715 | 14.39555047 | 0.01 | 0.01 |
-| Geranium molle | 119.11 | 141.13 | 3590 | 590 | 30.14020653 | 25.43753986 | 0.03 | 0.02 |
-| Alliaria petiolata | 810.5 | 892.12 | 17720 | 3630 | 21.8630475 | 19.86279873 | 0.02 | 0.02 |
-| Geum urbanum | 184.8 | 194.62 | 2390 | 860 | 12.93290043 | 12.28034118 | 0.01 | 0.01 |
-| Ballota nigra | 105.88 | 110.4 | 2090 | 610 | 19.73932754 | 18.93115942 | 0.01 | 0.01 |
-| Geranium sanguineum | 82.62 | 97.56 | 1910 | 530 | 23.11788913 | 19.57769578 | 0.02 | 0.01 |
-| Acer platanoides | 1142.2 | 1343.79 | 15110 | 4530 | 13.22885659 | 11.24431645 | 0.01 | 0.01 |
-
-: Leaf properties: leaf area, leaf fresh and dry weight, LMA (leaf mass per area) and EWT (equivalent water thickness) {#tbl-data-summary}
-
-{{< pagebreak >}}
-
-| Species | SPAD mean | SPAD stdv | chl Eq. 1 | chl Eq. 2 | chl Eq. 3 | chl Eq. 4 | chl Eq. 5 |
-|---|---|---|---|---|---|---|---|
-| Philadelphus Coronarius | 47.65 | 1.681434573 | 265.467675 | 621.573065 | 608.2915624 | 61.56568402 | 62.75488428 |
-| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 243.8084 | 584.63104 | 564.7495158 | 56.03685716 | 56.72252579 |
-| Geranium molle | 36.1 | 3.919183588 | 155.5059 | 425.94494 | 386.0649413 | 36.23261347 | 35.51025089 |
-| Alliaria petiolata | 27.64 | 1.416725489 | 93.583776 | 301.9521344 | 257.008503 | 24.57293509 | 23.40017292 |
-| Geum urbanum | 28.03 | 2.746330886 | 96.092139 | 307.3093226 | 262.3688604 | 25.01677595 | 23.85444116 |
-| Ballota nigra | 38.44 | 4.576558629 | 175.413216 | 463.1218304 | 426.6499938 | 40.34089174 | 39.8524312 |
-| Geranium sanguineum | 48.9 | 6.534332407 | 279.1283 | 644.56894 | 635.7441646 | 65.20131502 | 66.7438002 |
-| Acer platanoides | 31.59 | 5.872998477 | 120.535811 | 357.8139034 | 313.8933755 | 29.4575616 | 28.43095204 |
-
-: Leaf properties: SPAD measurements and chlorophyll concentration from 5 different equations {#tbl-data-summary}
-
-{{< pagebreak >}}
-
-<!-- change back to portrait -->
-\KOMAoptions{paper=portrait,pagesize} 
-\recalctypearea
-
-Material used:
-
-- Smartphone (GPS + plant species determination)
-- Book (Flora Vegetativa)
-- Survey sheet
-
-{{< pagebreak >}}
-
 # Task 2: Leaf chlorophyll measurements
@@
 {{< pagebreak >}}
-
-## Leaf images
-
-{{< pagebreak >}}
-
-## Recorded species tables
-
-{{< pagebreak >}}
-
-<!--
-Table in landscape format (uncomment if needed)
-
-# landscape table format (better?)
-
-::::::::{.landscape data-latex=""}
-
-\begin{table}[]
-\fontsize{22}{\baselineskip}\selectfont
-\renewcommand{\arraystretch}{4}
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
-\hline
-\textbf{Group-ID} &
-  \textbf{Plot-ID} &
-  \textbf{Subplot-ID} &
-  \textbf{Date} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}X-Coordinate \\ center\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Y-Coordinate \\ center\end{tabular}} &
-  \textbf{Litter / \%} &
-  \textbf{Soil / \%} &
-  \textbf{Water / \%} &
-  \textbf{Moss / \%} &
-  \textbf{Height / cm} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Acer \\ platanoides\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Achillea \\ millefolium\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Bellis \\ perennis\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Capsella \\ bursa-pastoris\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Dactylis \\ glomerata\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Elymus \\ repens\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ molle\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Geranium \\ pusillum\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Lolium \\ perenne\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Poa \\ annua\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Polygonum \\ aviculare\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Trifolium \\ repens\end{tabular}} &
-  \textbf{\begin{tabular}[c]{@{}c@{}}Veronica \\ arvensis\end{tabular}} \\ \hline
-4 & 1 & 1 & 16.05.2023 & 51.332222 & 12.384167 & 1.5 & 11 & 0 & 8 & 6 &     &     & 15 & 4 &   & 2  & 15 & 1 & 50 & 10 & 0.5 & 10 & 1   \\ \hline
-4 & 1 & 2 & 16.05.2023 &           &           & 1   & 4  & 0 & 1 & 6 &     & 0.5 & 22 &   & 2 &    & 18 & 1 & 55 & 15 & 0.5 &    & 1   \\ \hline
-4 & 1 & 3 & 16.05.2023 &           &           & 1   & 3  & 0 & 2 & 6 & 0.5 & 3   & 8  &   &   & 1  & 19 &   & 38 & 19 &     & 18 & 0.5 \\ \hline
-4 & 1 & 4 & 16.05.2023 &           &           & 1   & 10 & 0 & 1 & 4 &     &     & 6  &   &   & 15 & 17 &   & 28 & 10 & 0.5 & 1  & 2   \\ \hline
-\end{tabular}%
-}
-\caption{Subplot x species table for the recorded herbaceous vegetation. The values are the cover fraction of the total plot in {[}\%{]}.}
-\end{table}
-::::::::
-
--->
 
@@
 Finally we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Aria (LMA).
+
+# Task 3 & 5: Leaf water content
+
+<!-- Task 3 and 5 -->
+The total fresh weight and dry weight of the leave samples were used to determine the equivalent water thickness (EWT, @tbl-data-summary).
+The total weight of the leaves was determined through carefully placing the samples on precision laboratory balances.
+The equivalent water thickness is calculated by substracting the weight of the dry sample from the weight of the fresh sample and then deviding by the product of the density of water and the leaf area (@eq-EWT).
+
+ $$
+ EWT = \frac{fresh\: weight\: [mg]-dry\: weight\: [mg]}{1000\:\frac{mg}{cm^{3}}\; \; leaf\: area\: [cm^{2}]}
+ $$ {#eq-EWT}
+
+# Appendix
+
+<!-- change to landscape -->
+\KOMAoptions{paper=landscape,pagesize} 
+\recalctypearea
+
+<!-- please feel free to make these tables look prettier-->
+
+| Species | Leaf area (scanner) cm² | Leaf area (camera) cm² | Fresh weight [mg] | Dry weight [mg] | LMA [mg/cm²] scanner | LMA [mg/cm²] camera | EWT [mg/cm²] scanner | EWT [mg/cm²] camera |
+|---|---|---|---|---|---|---|---|---|
+| Philadelphus Coronarius | 719.97 | 751.19 | 13830 | 2440 | 19.20913371 | 18.41078822 | 0.02 | 0.02 |
+| Spirae Chamaedryfolia | 128.27 | 122.26 | 1760 | 620 | 13.72105715 | 14.39555047 | 0.01 | 0.01 |
+| Geranium molle | 119.11 | 141.13 | 3590 | 590 | 30.14020653 | 25.43753986 | 0.03 | 0.02 |
+| Alliaria petiolata | 810.5 | 892.12 | 17720 | 3630 | 21.8630475 | 19.86279873 | 0.02 | 0.02 |
+| Geum urbanum | 184.8 | 194.62 | 2390 | 860 | 12.93290043 | 12.28034118 | 0.01 | 0.01 |
+| Ballota nigra | 105.88 | 110.4 | 2090 | 610 | 19.73932754 | 18.93115942 | 0.01 | 0.01 |
+| Geranium sanguineum | 82.62 | 97.56 | 1910 | 530 | 23.11788913 | 19.57769578 | 0.02 | 0.01 |
+| Acer platanoides | 1142.2 | 1343.79 | 15110 | 4530 | 13.22885659 | 11.24431645 | 0.01 | 0.01 |
+
+: Leaf properties: leaf area, leaf fresh and dry weight, LMA (leaf mass per area) and EWT (equivalent water thickness) {#tbl-data-summary}
+
+{{< pagebreak >}}
+
+| Species | SPAD mean | SPAD stdv | chl Eq. 1 | chl Eq. 2 | chl Eq. 3 | chl Eq. 4 | chl Eq. 5 |
+|---|---|---|---|---|---|---|---|
+| Philadelphus Coronarius | 47.65 | 1.681434573 | 265.467675 | 621.573065 | 608.2915624 | 61.56568402 | 62.75488428 |
+| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 243.8084 | 584.63104 | 564.7495158 | 56.03685716 | 56.72252579 |
+| Geranium molle | 36.1 | 3.919183588 | 155.5059 | 425.94494 | 386.0649413 | 36.23261347 | 35.51025089 |
+| Alliaria petiolata | 27.64 | 1.416725489 | 93.583776 | 301.9521344 | 257.008503 | 24.57293509 | 23.40017292 |
+| Geum urbanum | 28.03 | 2.746330886 | 96.092139 | 307.3093226 | 262.3688604 | 25.01677595 | 23.85444116 |
+| Ballota nigra | 38.44 | 4.576558629 | 175.413216 | 463.1218304 | 426.6499938 | 40.34089174 | 39.8524312 |
+| Geranium sanguineum | 48.9 | 6.534332407 | 279.1283 | 644.56894 | 635.7441646 | 65.20131502 | 66.7438002 |
+| Acer platanoides | 31.59 | 5.872998477 | 120.535811 | 357.8139034 | 313.8933755 | 29.4575616 | 28.43095204 |
+
+: Leaf properties: SPAD measurements and chlorophyll concentration from 5 different equations {#tbl-data-summary}
+
+{{< pagebreak >}}
+
+\KOMAoptions{paper=portrait,pagesize} 
+\recalctypearea

### 2023-06-13 submission_2/submission_2.qmd
changed data table

@@
 |---|---|---|---|---|---|---|---|
-| Philadelphus Coronarius | 47.65 | 1.681434573 | 265.467675 | 621.573065 | 608.2915624 | 61.56568402 | 62.75488428 |
-| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 243.8084 | 584.63104 | 564.7495158 | 56.03685716 | 56.72252579 |
-| Geranium molle | 36.1 | 3.919183588 | 155.5059 | 425.94494 | 386.0649413 | 36.23261347 | 35.51025089 |
-| Alliaria petiolata | 27.64 | 1.416725489 | 93.583776 | 301.9521344 | 257.008503 | 24.57293509 | 23.40017292 |
-| Geum urbanum | 28.03 | 2.746330886 | 96.092139 | 307.3093226 | 262.3688604 | 25.01677595 | 23.85444116 |
-| Ballota nigra | 38.44 | 4.576558629 | 175.413216 | 463.1218304 | 426.6499938 | 40.34089174 | 39.8524312 |
-| Geranium sanguineum | 48.9 | 6.534332407 | 279.1283 | 644.56894 | 635.7441646 | 65.20131502 | 66.7438002 |
-| Acer platanoides | 31.59 | 5.872998477 | 120.535811 | 357.8139034 | 313.8933755 | 29.4575616 | 28.43095204 |
+| Philadelphus Coronarius | 47.65 | 1.681434573 | 26.5467675 | 55.53686838 | 54.35018076 | 61.56568402 | 62.75488428 |
+| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 24.38084 | 52.23613916 | 50.45974688 | 56.03685716 | 56.72252579 |
+| Geranium molle | 36.1 | 3.919183588 | 15.55059 | 38.057711 | 34.49447706 | 36.23261347 | 35.51025089 |
+| Alliaria petiolata | 27.64 | 1.416725489 | 9.3583776 | 26.97909046 | 22.96342652 | 24.57293509 | 23.40017292 |
+| Geum urbanum | 28.03 | 2.746330886 | 9.6092139 | 27.45774932 | 23.44236855 | 25.01677595 | 23.85444116 |
+| Ballota nigra | 38.44 | 4.576558629 | 17.5413216 | 41.37942519 | 38.12070678 | 40.34089174 | 39.8524312 |
+| Geranium sanguineum | 48.9 | 6.534332407 | 27.91283 | 57.59152447 | 56.80304051 | 65.20131502 | 66.7438002 |
+| Acer platanoides | 31.59 | 5.872998477 | 12.0535811 | 31.97027796 | 28.04602719 | 29.4575616 | 28.43095204 |
 

### 2023-06-13 submission_2/submission_2.qmd
add description of varying equations (chlorophyll)

@@
 
+It is to mention that the unit of the equation from Monje & Bugbee (1992) was wrong and therefore we divided the resulting values by 10.
+We further recalculated the equations from Markwell et al.
+(1995) from $\mu mol/mm^2$ to $\mu g/mm^2$ by multiplicating the values with 0.089348898 (because the molar mass of chlorophyll A is 893.48898 g/mol).
+
 The results of the conversions can be seen in the attached data table as well as in the figure below.
@@
 ![Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.](SPAD_conversion_variation.png "Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.")
+
+All equations yield relatively similar results (see curves 2-5), except the one from Monje & Bugbee (1992), which results in lower values.
 

### 2023-06-13 submission_2/submission_2.qmd
added text for comparison to task4

--- 2023-06-13 submission_2/submission_2.qmd
@@
 
+## Comparison to LEDA Database
+
+Another task was to compare our LMA values with those from the LEDA database.
+There, the values were only available as SLA.
+So we had to convert them into LMA first.
+The result can be seen in the plots in the appendix.
+If several values were available for one plant, we calculated the mean value.
+For two plants no values were available in the LEDA database.
+Overall, our values are slightly higher than those in the database, but are very similar overall.
+
 # Task 3 & 5: Leaf water content

### 2023-06-13 submission_2/submission_2.qmd
Update submission_2.qmd

--- 2023-06-13 submission_2/submission_2.qmd
@@
 Finally we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Aria (LMA).
+## Comparison to LEDA Database
+
+Another task was to compare our LMA values with those from the LEDA database.
+There, the values were only available as SLA.
+So we had to convert them into LMA first.
+The result can be seen in the plots in the appendix.
+If several values were available for one plant, we calculated the mean value.
+For two plants no values were available in the LEDA database.
+Overall, our values are slightly higher than those in the database, but are very similar overall.
 

### 2023-06-13 submission_2/submission_2.qmd
add scan vs cam plot
Updates report & new images path

@@
 
-![Table with formulas to convert measured SPAD pigment values to chlorophyll per area values.](SPAD_conversion_table.png "Table with formulas to convert measured SPAD pigment values to chlorophyll per area values.")
+![Table with formulas to convert measured SPAD pigment values to chlorophyll per area values.](imgs/SPAD_conversion_table.png "Table with formulas to convert measured SPAD pigment values to chlorophyll per area values.")
 
@@
 
-![Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.](SPAD_conversion_variation.png "Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.")
+![Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.](imgs/SPAD_conversion_variation.png "Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values."){width=80%}
 
@@
 
+![Scatterplot comparing results of Leaf area from the scanner vs Leaf area out of camera photos.](imgs/leaf-area_scan_vs_cam.png "Scatterplot comparing results of Leaf area from the scanner vs Leaf area out of camera photos."){width=70%}
+
 ## LMA
@@
 Finally we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Aria (LMA).
+
 ## Comparison to LEDA Database
 
-Another task was to compare our LMA values with those from the LEDA database.
-There, the values were only available as SLA.
-So we had to convert them into LMA first.
-The result can be seen in the plots in the appendix.
+Another task was to compare our LMA values with those from the LEDA database (See Figure 2 in this section).
+There, the values were available as mean SLA, so we converted them into LMA first.
 If several values were available for one plant, we calculated the mean value.
@@
 Overall, our values are slightly higher than those in the database, but are very similar overall.
+
+![Comparison of our LMA values (red) to values from the LEDA database (blue)](imgs/leda_lma_comparison.png)
 
@@
  $$ {#eq-EWT}
+{{< pagebreak >}}
 

### 2023-06-13 submission_2/submission_2.qmd
better figure rendering
update tabel, correct LEDA values fig2

@@
 subsubtitle: "Leaf traits measurements"
-author: 
-    - "Group 4" 
-    - "Denis Streitmatter"
-    - "Marie-Louise Korte"
-    - "Louis Trinkle"
-    - "Maxwell Sivertsen"
-    - "Fabian Götz"
-    - "Lukas Erzfeld"
+author: | 
+    | **Group 4** 
+    | Denis Streitmatter 
+    | Marie-Louise Korte 
+    | Louis Trinkle 
+    | Maxwell Sivertsen 
+    | Fabian Götz 
+    | Lukas Erzfeld
 date: last-modified
@@
 
+{{< pagebreak >}}
+
 ## LMA
@@
 
+{{< pagebreak >}}
+
 # Appendix
-
 <!-- change to landscape -->
@@
 \recalctypearea
+<!-- please feel free to make these tables look prettier-->
+## Table 1: Leaf area and weight
+| Species | Leaf area [cm²] scanner | Leaf area [cm²] camera | Fresh weight [mg] | Dry weight [mg] | SLA [cm²/mg] | EWT [mg/cm²] scanner | EWT [mg/cm²] camera |
+|---|---|---|---|---|---|---|---|
+| Philadelphus Coronarius | 719.97 | 751.19 | 13830 | 2440 | 0.31 | 0.02 | 0.02 |
+| Spirae Chamaedryfolia | 128.27 | 122.26 | 1760 | 620 | 0.20 | 0.01 | 0.01 |
+| Geranium molle | 119.11 | 141.13 | 3590 | 590 | 0.24 | 0.03 | 0.02 |
+| Alliaria petiolata | 810.50 | 892.12 | 17720 | 3630 | 4.07 | 0.02 | 0.02 |
+| Geum urbanum | 184.80 | 194.62 | 2390 | 860 | 0.23 | 0.01 | 0.01 |
+| Ballota nigra | 105.88 | 110.40 | 2090 | 610 | 0.18 | 0.01 | 0.01 |
+| Geranium sanguineum | 82.62 | 97.56 | 1910 | 530 | 5.43 | 0.02 | 0.01 |
+| Acer platanoides | 1142.20 | 1343.79 | 15110 | 4530 | 3.37 | 0.01 | 0.01 |
 
-<!-- please feel free to make these tables look prettier-->
-
-| Species | Leaf area (scanner) cm² | Leaf area (camera) cm² | Fresh weight [mg] | Dry weight [mg] | LMA [mg/cm²] scanner | LMA [mg/cm²] camera | EWT [mg/cm²] scanner | EWT [mg/cm²] camera |
-|---|---|---|---|---|---|---|---|---|
-| Philadelphus Coronarius | 719.97 | 751.19 | 13830 | 2440 | 19.20913371 | 18.41078822 | 0.02 | 0.02 |
-| Spirae Chamaedryfolia | 128.27 | 122.26 | 1760 | 620 | 13.72105715 | 14.39555047 | 0.01 | 0.01 |
-| Geranium molle | 119.11 | 141.13 | 3590 | 590 | 30.14020653 | 25.43753986 | 0.03 | 0.02 |
-| Alliaria petiolata | 810.5 | 892.12 | 17720 | 3630 | 21.8630475 | 19.86279873 | 0.02 | 0.02 |
-| Geum urbanum | 184.8 | 194.62 | 2390 | 860 | 12.93290043 | 12.28034118 | 0.01 | 0.01 |
-| Ballota nigra | 105.88 | 110.4 | 2090 | 610 | 19.73932754 | 18.93115942 | 0.01 | 0.01 |
-| Geranium sanguineum | 82.62 | 97.56 | 1910 | 530 | 23.11788913 | 19.57769578 | 0.02 | 0.01 |
-| Acer platanoides | 1142.2 | 1343.79 | 15110 | 4530 | 13.22885659 | 11.24431645 | 0.01 | 0.01 |
-
-: Leaf properties: leaf area, leaf fresh and dry weight, LMA (leaf mass per area) and EWT (equivalent water thickness) {#tbl-data-summary}
+: Leaf properties: leaf area, leaf fresh and dry weight, SLA (surface leaf area) and EWT (equivalent water thickness) {#tbl-data-summary}
 
@@
 
-| Species | SPAD mean | SPAD stdv | chl Eq. 1 | chl Eq. 2 | chl Eq. 3 | chl Eq. 4 | chl Eq. 5 |
+## Table 2: Leaf mass per area
+| Species | Mean LEDA Database LMA [mg/cm²]  | Leaf mass per area [mg/cm²] scanner | Leaf mass per area [mg/cm²] camera |
+|---|---|---|---|
+| Philadelphus Coronarius |  | 3.39 | 3.25 |
+| Spirae Chamaedryfolia |  | 4.83 | 5.07 |
+| Geranium molle | 3.62 | 4.95 | 4.18 |
+| Alliaria petiolata | 2.45 | 4.48 | 4.07 |
+| Geum urbanum | 2.65 | 4.65 | 4.42 |
+| Ballota nigra | 4.71 | 5.76 | 5.53 |
+| Geranium sanguineum | 4.58 | 6.41 | 5.43 |
+| Acer platanoides | 5.09 | 3.97 | 3.37 |
+
+: LMA (leaf mass per area) from the LEDA database compared to our data {#tbl-data-LMA}
+
+{{< pagebreak >}}
+
+## Table 3: SPAD measurements and chlorophyll
+| Species | SPAD mean | SPAD stdv | Chl Eq. 1 | Chl Eq. 2 | Chl Eq. 3 | Chl Eq. 4 | Chl Eq. 5 |
 |---|---|---|---|---|---|---|---|
-| Philadelphus Coronarius | 47.65 | 1.681434573 | 26.5467675 | 55.53686838 | 54.35018076 | 61.56568402 | 62.75488428 |
-| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 24.38084 | 52.23613916 | 50.45974688 | 56.03685716 | 56.72252579 |
-| Geranium molle | 36.1 | 3.919183588 | 15.55059 | 38.057711 | 34.49447706 | 36.23261347 | 35.51025089 |
-| Alliaria petiolata | 27.64 | 1.416725489 | 9.3583776 | 26.97909046 | 22.96342652 | 24.57293509 | 23.40017292 |
-| Geum urbanum | 28.03 | 2.746330886 | 9.6092139 | 27.45774932 | 23.44236855 | 25.01677595 | 23.85444116 |
-| Ballota nigra | 38.44 | 4.576558629 | 17.5413216 | 41.37942519 | 38.12070678 | 40.34089174 | 39.8524312 |
-| Geranium sanguineum | 48.9 | 6.534332407 | 27.91283 | 57.59152447 | 56.80304051 | 65.20131502 | 66.7438002 |
-| Acer platanoides | 31.59 | 5.872998477 | 12.0535811 | 31.97027796 | 28.04602719 | 29.4575616 | 28.43095204 |
+| Philadelphus Coronarius | 47.65 | 1.68 | 26.55 | 55.54 | 54.35 | 61.57 | 62.75 |
+| Spirae Chamaedryfolia | 45.6 | 3.22 | 24.38 | 52.24 | 50.46 | 56.04 | 56.72 |
+| Geranium molle | 36.1 | 3.92 | 15.55 | 38.06 | 34.49 | 36.23 | 35.51 |
+| Alliaria petiolata | 27.64 | 1.42 | 9.36 | 26.98 | 22.96 | 24.57 | 23.40 |
+| Geum urbanum | 28.03 | 2.75 | 9.61 | 27.46 | 23.44 | 25.02 | 23.85 |
+| Ballota nigra | 38.44 | 4.58 | 17.54 | 41.38 | 38.12 | 40.34 | 39.85 |
+| Geranium sanguineum | 48.9 | 6.53 | 27.91 | 57.59 | 56.80 | 65.20 | 66.74 |
+| Acer platanoides | 31.59 | 5.87 | 12.05 | 31.97 | 28.05 | 29.46 | 28.43 |
 
-: Leaf properties: SPAD measurements and chlorophyll concentration from 5 different equations {#tbl-data-summary}
+: Leaf properties: SPAD measurements and chlorophyll concentration from 5 different equations {#tbl-data-chl}
 

### 2023-06-13 submission_2/submission_2.qmd
Update submission_2.qmd

@@
 The plant species that were sampled were choosen based on their abundance in the park.
+The plants were identified using the smartphone app 'FloraIncognita' and the book 'Flora Vegetativa - Ein Bestimmungsbuch für Pflanzen der Schweiz im blütenlosen Zustand' (2020) by Eggenberg and Möhl and the botanical names of the species were recorded in a table.
+There were slight uncertainties distinguishing between *Geranium molle* and *G.
+sanguineum*, all remaining species were clearly identifiabe.
+
 For each species 10 leaves from different individuals were sampled.
-When sampling leaves we were careful to sample the whole leaf.
+When sampling leaves, we were careful to sample the whole leaf including petiole and all other leaf attachments.
 Leaves were chosen based on the following criteria, healthy, fully grown and from a sun-lit environment where possible.
 We sampled trees, shrubs and herbs.
-The plants were identified using the smartphone app 'FloraIncognita' and the latin names of the species were recorded in a table.
-The leaves were collected in labeled paper bags and taken back to the seminar room for further analyisis.
+The leaves were collected in labeled paper bags and taken back to the seminar room for further analyses.
+
 The leaf traits recorded for each species were fresh weight, dry weight, leaf area, and chlorophyll content.
 The fresh weight was determined first using a precision balance.
-In the next step chlorophyll concentration was derived from plant greenness using a Soil Plant Analysis Development (SPAD) meter.
-We used the leaf scanner to measure leaf area, we also took pictures of the leafs on an A4 sheet to derive leaf area using a script for image analysis.
+In the next step, chlorophyll concentration was derived from plant greenness using a Soil Plant Analysis Development (SPAD) meter (for more detailed information see Task 2).
+We used the leaf scanner to measure leaf area, and we took pictures of the leaves on a white A4 sheet to derive leaf area using a script for image analysis (see Task 4).
 The leaves were then put in a drying cabinate at 50° to 55° for 48h.
-In a last step dry weight was determined.
-All the data was recorded in a google sheet.
+In a last step the dry weight was determined.
+All the data were recorded in a google sheet and can be seen in the tables in the Appendix.
 
@@
 sanguineum*) to six spots (*Acer platanoides, Alliaria petiolata, Ballota nigra, Geum urbanum, Philadelphus Coronarius, Spirae Chamaedryfolia*), depending on the leaf size and shape.
-During the measurement, we padi attention so that no obvious outliers were recorded, no veins were measured and that the optical window of the instrument was fully covered by the leaf.
-For each of the ten sampled leaves, the mean of the four to six measurements was written down on the leaf trait sheet (see columns *SPAD1...SPAD10* of table < INSERT TABLE NUMBERING>).
-Afterwards, the mean and standard deviation for all the leaves of each species were calculated.
+During the measurement, we paid attention so that no obvious outliers were recorded, no veins were measured and that the optical window of the instrument was fully covered by the leaf.
+For each of the ten sampled leaves, the mean of the four to six measurements was written down on the leaf trait sheet.
+Afterwards, the mean and standard deviation for all the leaves of each species were calculated (see Table 3).
 
@@
 We further recalculated the equations from Markwell et al.
-(1995) from $\mu mol/mm^2$ to $\mu g/mm^2$ by multiplicating the values with 0.089348898 (because the molar mass of chlorophyll A is 893.48898 g/mol).
+(1995) from $\mu mol/m^2$ to $\mu g/cm^2$ by multiplicating the values with 0.089348898 (because the molar mass of chlorophyll A is 893.48898 g/mol and the area unit of the equations was $m^2$).
 
-The results of the conversions can be seen in the attached data table as well as in the figure below.
+The results of the conversions can be seen in the attached data Table 3 (columns *Chl Eq.
+1* ...
+*Chl Eq.
+5*) as well as in Figure 2 below.
+All equations yield relatively similar results (see curves 2-5), except the one from Monje & Bugbee (1992), which results in lower chlorophyll A values.
 
-![Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.](imgs/SPAD_conversion_variation.png "Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values."){width=80%}
-
-All equations yield relatively similar results (see curves 2-5), except the one from Monje & Bugbee (1992), which results in lower values.
+![Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values.
+The unit of the chlorophyll values is $\mu g/cm^2$.](imgs/SPAD_conversion_variation.png "Variation of the different equations used to convert the measured SPAD-values to real chlorophyll values."){width=80%}
 
@@
 In the next step, we determined the mass per leaf area.
-For this task two different methods were used: A specialized scanner for this exact purpose and a python script calculating the leaf area based of photos on an a DIN-A4 paper.
+For this task two different methods were used: A specialized scanner for this exact purpose and a python script calculating the leaf area based of photos on a white DIN-A4 paper.
 
@@
 
- We took the sheets out of the paper bag and placed them one by one into the sheet scanner.
-We had to make sure that the sheets were fully spread out, did not overlap, and were within the measurement range area.
-For some sheets, the scanner was not large enough to measure all 10 sheets at once.
-So we did multiple scans and added up and documented the measured leaf area for each species.
-
-To check the correctness of the scanner we performed a second type of leaf surface measurement.
-For this, we placed the leaves on a DinA4 sheet and took a photo.
-We wrote a small script to convert the scans of sheets and paper into black and white images and calculate the cumulative sheet area from the proportion of black pixels of the total paper area.
-Then we compared the measured area with the result of the sheet scanner.
-They were very similar everywhere with slight variations (see documentation).
-There were two different approaches for measuring the leaf area: Using a specific tool or calculating it on the base of photos of the leafs on standardised A4 paper sheets
+We took the leaves out of the paper bag and placed them one by one into the sheet scanner.
+We made sure that the sheets were fully spread out, did not overlap, and were within the measurement range area.
+For some leaves, the scanner was not large enough to measure all 10 pieces at once.
+Therefore we conducted multiple scans, added the values up and documented the measured leaf area for each species.
 
@@
 
-For this approach we took Photos of single or multiple leafs on one (or two) A4 papers. 
-These photos were then preprocessed with the image-software `gimp` and then calculated the leaf size with a [python script](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/leaf_area_measurement.py).
-The photos (unprocessed and otherwise) can be found [here](https://github.com/yum-yab/ground_truthing_module/tree/main/leaf_area_measure/leaf_images)
+To validate the accuracy of the scanner, we performed a second type of leaf surface measurement.
+For this approach, we placed single or multiple leaves on one (or two) white DIN-A4 papers and took a photo, covering all sampled leaves of each species.
+We preprocessed these photos using the free image-software `gimp` and then calculated the leaf size with this [python script](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/leaf_area_measurement.py).
+The photos (unprocessed and otherwise) can be found [here](https://github.com/yum-yab/ground_truthing_module/tree/main/leaf_area_measure/leaf_images).
 
@@
 
-The goal of this step is to reduce the image to the necessary parts, mainly cutting out the A4 paper sheet. For this a few steps were performed:
+The goal of this step was to reduce the image to the necessary parts, mainly cutting out the A4 paper sheet. For this, these steps were performed:
 
-1. Load the image in gimp
+1. Load the image in `gimp`
 2. Use the `lasso-select tool` for selecting the corners of the A4 paper
@@
 5. Use `Image > Crop to content` to reduce the image size
-    - this step is optional since alpha channel (= transparency) is ignored in the script, but this reduces file size and therefor makes it a bit faster
+    - this step is optional since alpha channel (= transparency) is ignored in the script, but this reduces file size and therefore makes it a bit faster
 6. Save images in proper format for script:
@@
 The code and the description how to use it can be found in [this git repository](https://github.com/yum-yab/ground_truthing_module/blob/main/leaf_area_measure/). 
-Here we will only describe the main ideas of the process. 
-Since we know the area of the A4 paper, to determine leaf size we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` ($\equiv$ DIN-A4) and therefor we can calculate the area of the leaf.
-Unfortunately shadows of not fully flat leaves can add a lot of noise to the trivial approach (converting it to greyscale and cutting at a certain threshold). 
-To fix this we transposed the image in the `HSV` color space, where all grey/white colors have a very low saturation value (compared to the green leaves), so we determine a  saturation threshold here (a comparison feature between Bitmask, `HSV` and original image was implemented to find a fitting value) and generate the bitmask in this way.
+Here, we will only describe the main ideas of the process. 
+Since we know the area of the A4 paper, to determine leaf size we only need to know the number of `leaf-pixels` in relation to `leaf-pixels + paper-pixels` ($\equiv$ DIN-A4) and therefore we can calculate the area of the leaf.
+
+Unfortunately, shadows of not fully flat leaves can add a lot of noise to the trivial approach (converting it to greyscale and cutting at a certain threshold). 
+To fix this, we transposed the image into the `HSV` color space, where all grey/white colors have a very low saturation value (compared to the green leaves), determine a  saturation threshold here (a comparison feature between Bitmask, `HSV` and original image was implemented to find a fitting value) and generate the bitmask in this way.
 The transparent parts are not considered during the area calculation, only the white (paper), grey (shadows, counted as paper) and green (leaf) pixels.
+
+The output of the script for each plant species can be seen in Table 1.
+Then, we compared the measured area with the result of the sheet scanner.
+They showed similar values with slight variations (see Table 1 and Figure 1: the red line denotes the line of perfect fit, the blue dots are the respective values of the camera/scanner approaches).
 
@@
 
-To calculate the biomass of the sheets, we now had to determine the dry weight of the sheets.
-To do this, we put the leaves back in the paper bag, and put them in the drying oven at 50 - 55 °C.
-The samples were dried for 72h.
-Then we measured the dry weight and noted the values.
-Finally we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Aria (LMA).
+To calculate the biomass of the leaves, we determined their dry weight.
+To do this, we put the samples back in the paper bag and afterwards in the drying oven at 50 - 55 °C for 72h.
+Then, we measured the dry weight using a precision balance and noted the values.
+Finally, we calculated the ratio of dry weight to leaf area in mg/cm², which is the Leaf mass per Area (LMA).
 
@@
 
-Another task was to compare our LMA values with those from the LEDA database (See Figure 2 in this section).
-There, the values were available as mean SLA, so we converted them into LMA first.
-If several values were available for one plant, we calculated the mean value.
-For two plants no values were available in the LEDA database.
-Overall, our values are slightly higher than those in the database, but are very similar overall.
+We further compared our LMA values with those from the LEDA database (See Figure 2 in this section).
+There, the values were available as mean SLA, so we converted them into LMA first (LMA = \frac{1}{SLA}).
+If several values were available for one plant species, we calculated the mean value.
+For two species (*Philadelphus Coronarius* and *Spirae Chamaedryfolia*) no values were available in the LEDA database.
+Overall, our values are slightly higher than those in the database, but generally similar (see Table 2 and Figure 2).
+Reasons for this may be inaccuracies in our sheet size script or inaccuracies in the sampling of the leaves by the respective processor (our group or people recorded in the LEDA database).
 
@@
 <!-- Task 3 and 5 -->
-The total fresh weight and dry weight of the leave samples were used to determine the equivalent water thickness (EWT, @tbl-data-summary).
+The total fresh weight and dry weight of the leaf samples were used to determine the equivalent water thickness (EWT, @tbl-data-summary).
 The total weight of the leaves was determined through carefully placing the samples on precision laboratory balances.
-The equivalent water thickness is calculated by substracting the weight of the dry sample from the weight of the fresh sample and then deviding by the product of the density of water and the leaf area (@eq-EWT).
+The equivalent water thickness is calculated by substracting the weight of the dry sample from the weight of the fresh sample and then dividing by the product of the density of water and the leaf area (@eq-EWT).
 

### 2023-07-18 submission_2/submission_2.qmd
Update submission_2.qmd

@@
 We further compared our LMA values with those from the LEDA database (See Figure 2 in this section).
-There, the values were available as mean SLA, so we converted them into LMA first (LMA = \frac{1}{SLA}).
+There, the values were available as mean SLA, so we converted them into LMA first ($LMA = \frac{1}{SLA}$).
 If several values were available for one plant species, we calculated the mean value.
@@
 ![Comparison of our LMA values (red) to values from the LEDA database (blue)](imgs/leda_lma_comparison.png)
+
+## Comparison to LEDA Database
+
+Another task was to compare our LMA values with those from the LEDA database.
+There, the values were only available as SLA.
+So we had to convert them into LMA first.
+The result can be seen in the plots in the appendix.
+If several values were available for one plant, we calculated the mean value.
+For two plants no values were available in the LEDA database.
+Overall, our values are slightly higher than those in the database, but are very similar overall.
 

### 2023-07-18 submission_4/submission_4.qmd

---
title: "submission_4"
subtitle: "Documentation"
subsubtitle: "Leaf traits measurements"
author: | 
    | **Group 4** 
    | Denis Streitmatter 
    | Marie-Louise Korte 
    | Louis Trinkle 
    | Maxwell Sivertsen 
    | Fabian Götz 
    | Lukas Erzfeld
date: last-modified
format:
    pdf:
        documentclass: scrreport
        theme: united
        toc: true
        toc-depth: 2
        number-sections: false 
        number-depth: 1
        toc-title: Table of Contents
        toc-location: body
        colorlinks: true
header-includes:  
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{{< pagebreak >}}

#

### 2023-07-19 submission_4/submission_4.qmd
Adds sources for subission 4

@@
 
-#
+# Submission 4: TLS and Sentinel 2 comparison
+
+## General study concept and conditions
+
+The purpose of this document is the comparison of terrestial laser/LiDAR scanning and Sentinel 2 (satellite) data of the same spot.
+
+**Insert some description and the data of the plot. I wasn't there, so probably someone else should do it**
+
+## Algorithm idea
+
+## Results
+
+This section lists, describes and tries to explain the plots and graphs generated by the algorithm described in the previous section.
+
+### Tree height statistic comparison
+
+In this section the general measurements are compared.
+
+*First the general heigths:*
+
+![Height calculation of TLS (resampled to resolution of the sentinel data)](TLS_plots/figures/chm_height.png)
+
+### Residual computation
+
+In this section we compute the residuals of the

### 2023-07-19 submission_4/code/TLS_plots.R

@@ -1,0 +1,229 @@
+##Claudia Guimaraes-Steinicke
+## Second exercises on Ground truthing
+# June 27th 2023
+
+# Load the library
+library(devtools)
+library(canopyLazR)
+library(rLiDAR)
+library(forestr)
+library(lidR)
+library(rayshader)
+library(RCSF)
+library(rayshader)
+library(rgdal)
+library(sf)
+library(ggplot2)
+
+
+setwd("/home/denis/workspace/uni/ground_truthing/submission_4/code")
+
+
+### colorbar 
+col <- height.colors(25)
+
+### setting an extent to tls raster
+aoi <- shapefile("shapefile/polygon.shp")
+aoi
+
+
+### TLS canopy height model
+chm_IEP_new <- raster('chm_IEP.tif')
+chm_IEP <- crop(chm_IEP_new, aoi)
+plot(chm_IEP_new)
+
+
+## calculate standard deviation of chm_IEP
+h_tls_sd_new = aggregate(chm_IEP, fact = 10, fun = "sd")
+plot(h_tls_sd_new)
+hist(h_tls_sd_new)
+
+
+##############################################################################
+## Download the Sentinel 2 
+#############################################################################
+h_s2_top <- raster("ETH_GlobalCanopyHeight_DE_CHM_clip.tif")
+plot(h_s2_top)
+h_s2_top
+h_s2_sd <- raster("ETH_GlobalCanopyHeight_DE_STD_clip.tif")
+plot(h_s2_sd)
+
+
+# Resample chm_IEP to match the resolution and extent of Sentinel-2
+h_tls_top <- projectRaster(chm_IEP, h_s2_top, alignOnly = F)
+plot(h_tls_top)
+h_tls_sd <- projectRaster(h_tls_sd_new, h_s2_top, alignOnly = F)
+plot(h_tls_sd)
+
+##resample?
+h_tls_top_res <- resample(h_tls_top, h_s2_top, method = "bilinear")
+h_tls_sd_res <- resample(h_tls_sd, h_s2_top, method = "bilinear")
+
+## crop to area of interest
+h_s2_top <- crop(h_s2_top, aoi)
+h_s2_sd <- crop(h_s2_sd, aoi)
+
+### set the same extent
+plot(h_s2_top)
+hist(h_s2_top)
+
+plot(h_s2_sd)
+hist(h_s2_sd)
+
+
+################################################################################################
[…]

### 2023-07-19 submission_4/code/TLS_plots.R
Deletes double code repo

@@ -1,229 +1,0 @@
-##Claudia Guimaraes-Steinicke
-## Second exercises on Ground truthing
-# June 27th 2023
-
-# Load the library
-library(devtools)
-library(canopyLazR)
-library(rLiDAR)
-library(forestr)
-library(lidR)
-library(rayshader)
-library(RCSF)
-library(rayshader)
-library(rgdal)
-library(sf)
-library(ggplot2)
-
-
-setwd("/home/denis/workspace/uni/ground_truthing/submission_4/code")
-
-
-### colorbar 
-col <- height.colors(25)
-
-### setting an extent to tls raster
-aoi <- shapefile("shapefile/polygon.shp")
-aoi
-
-
-### TLS canopy height model
-chm_IEP_new <- raster('chm_IEP.tif')
-chm_IEP <- crop(chm_IEP_new, aoi)
-plot(chm_IEP_new)
-
-
-## calculate standard deviation of chm_IEP
-h_tls_sd_new = aggregate(chm_IEP, fact = 10, fun = "sd")
-plot(h_tls_sd_new)
-hist(h_tls_sd_new)
-
-
-##############################################################################
-## Download the Sentinel 2 
-#############################################################################
-h_s2_top <- raster("ETH_GlobalCanopyHeight_DE_CHM_clip.tif")
-plot(h_s2_top)
-h_s2_top
-h_s2_sd <- raster("ETH_GlobalCanopyHeight_DE_STD_clip.tif")
-plot(h_s2_sd)
-
-
-# Resample chm_IEP to match the resolution and extent of Sentinel-2
-h_tls_top <- projectRaster(chm_IEP, h_s2_top, alignOnly = F)
-plot(h_tls_top)
-h_tls_sd <- projectRaster(h_tls_sd_new, h_s2_top, alignOnly = F)
-plot(h_tls_sd)
-
-##resample?
-h_tls_top_res <- resample(h_tls_top, h_s2_top, method = "bilinear")
-h_tls_sd_res <- resample(h_tls_sd, h_s2_top, method = "bilinear")
-
-## crop to area of interest
-h_s2_top <- crop(h_s2_top, aoi)
-h_s2_sd <- crop(h_s2_sd, aoi)
-
-### set the same extent
-plot(h_s2_top)
-hist(h_s2_top)
-
-plot(h_s2_sd)
-hist(h_s2_sd)
-
-
-################################################################################################
[…]

### 2023-07-20 submission_3/UAV+S2.qmd
upated submission3

---
title: "UAV vs Sentinel-2 height comparisons"
subtitle: "Documentation"
author: "Fabian Götz"
date: 6/24/23
format:
    pdf:
        documentclass: scrreport
        theme: united
        toc: true
        toc-depth: 2
        number-sections: true
        number-depth: 1
        toc-title: Table of Contents
        toc-location: body
        colorlinks: true
header-includes:  
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{{< pagebreak >}}

# How well do Sentinel-based height estimates correspond to UAV-based height?

Date: 24.06.2023

Group-ID: 4

Participants: Denis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

Material used:

- ETH Sentinel-2 data
- UAV data

## Top canopy height per Sentinel-pixel

As the figures in Figure 1.1 - 1.3 show, the UAV estimates differ significantly from those of the S2 data. UAV peak height per pixel shows a scale with a maximum of 35 and a minimum of 0. In comparison, S2 peak height per pixel shows only a maximum of 33 and a minimum of 26. Thus, the range of UAV values is much larger. This suggests a much finer spatial resolution leading to the differences. This confirms what we have already learned, that UAV is able to record at a finer spatial resolution than satellites are able to, due to their distance. Thus, we assume that UAV are closer to ground truth than Sentinel-2. While the histograms show the general distribution of the data, the maps allow us to see the distributions based on the coordinates. In the map of the residuals it can be seen in another way that the measured values of the two measurement methods are different. Specifically, one can see that Sentinel-2 does not resolve accurately enough to detect locations with low tree heights. Thus, the values of Sentinel-2 are too high on average.

### Plot images

\begin{figure}[htbp]
  \centering

  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/uav_height.png}
    \caption{Map of UAV-based height estimation}
    \label{fig:image1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/uav_height_hist.png}
    \caption{Histogram of UAV-based height estimation}
    \label{fig:image2}
  \end{subfigure}

  \caption{UAV-based height estimation}
  \label{uav}
\end{figure}

\begin{figure}[htbp]

  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/s2_height.png}
    \caption{Map of Sentinel-2-based height estimation}
    \label{fig:image3}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/s2_height_hist.png}
    \caption{Histogram of Sentinel-2-based height}
    \label{fig:image4}
  \end{subfigure}

  \caption{Sentinel-2-based height estimation}
  \label{fig:sentinel}
\end{figure}

\begin{figure}[htbp]

  \includegraphics[width=\textwidth]{images/residuals.png}
  \caption{Residuals of the two height measurements}
  \label{fig:sentinel}
\end{figure}

{{< pagebreak >}}

## Standard deviation of height per Sentinel-pixel

The standard deviation (SD) of the UAV pixels ranges from 0 to 8 meters. In comparison, the S2 SD shows a height of the Sentinel pixel that ranges from about 6m at the low end to 7m as the upper value. This shows that UAV's values are much more spread out than Sentinel's. This indicates that UAV has a more accurate resolution of the data.

In summary, this exercise shows the differences between UAV and S2 data quality in terms of spatial resolution. The exercise shows the increase in available information when the spatial resolution is increased by using UAV.

### Plot images

\begin{figure}[htbp]

  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/uav_sd.png}
    \caption{Map of standard deviation of UAV measurement}
    \label{fig:image5}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/uav_sd_hist.png}
    \caption{Histogram of standard deviation of UAV measurement}
    \label{fig:image6}
  \end{subfigure}

  \caption{UAV standard deviation}
  \label{fig:uav_sd}
\end{figure}

\begin{figure}[htbp]

  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/s2_sd.png}
    \caption{Map of standard deviation of S2 measurement}
    \label{fig:image7}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/s2_sd_hist.png}
    \caption{Histogram of standard deviation of S2 measurement}
    \label{fig:image8}
  \end{subfigure}

  \caption{Sentinel-2 standard deviation}
  \label{fig:uav_sd}
\end{figure}

{{< pagebreak >}}

## Compare the Metashape DTM (DJI_202305311408_001_crane-2022-05_dem.tif) with the DTM based on the point cloud without the crane. Did the cleaning solve the problem?

The problem is not explicitly given here, however from the observations available it
appears as though the cleaning and creation of a DTM using the point clouds results in very low spatial resolution. This is what we would expect, given that the DTM created from the LiDAR data using the rasterization tool was set to a resolution of 10 - 30m. In our specific case it was set to 15.

### 2023-07-20 submission_3/old/uav_s2_height_comparison.r

@@ -1,0 +1,101 @@
+library(raster)
+
+# set your wd to where you have the downloaded data
+setwd()
+
+
+
+# Load the two raster files
+
+h_uav_dsm <- raster("uav/cloudcompare_dsm_v2.tif")
+
+crs(h_uav_dsm) <- ("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs")
+
+
+h_uav_dsm <- raster("uav/DJI_202305311408_001_crane-2022-05_dem.tif")
+
+# set 0 to NA
+
+h_uav_dsm <- reclassify(h_uav_dsm, cbind(0, NA))
+
+
+h_s2_top <- raster("sentinel_2_height/ETH_GlobalCanopyHeight_DE_CHM_clip.tif")
+
+h_s2_sd <- raster("sentinel_2_height/ETH_GlobalCanopyHeight_DE_STD_clip.tif")
+
+
+plot(h_uav_dsm-147)
+
+h_uav_top = h_uav_dsm - 147
+
+# calculate standard deviation per ~25m² pixels
+
+h_uav_sd = aggregate(h_uav_top, fact = 10, fun = "sd")
+
+
+# Resample raster1 to match the resolution and extent of raster2
+
+# projectRaster() - projects values of 1 raster object to a new object raster with another projection projectRaster(from, to, alignOnly)
+h_uav_top <- projectRaster(h_uav_top, h_s2_top, alignOnly = F)
+
+h_uav_sd <- projectRaster(h_uav_sd, h_s2_top, alignOnly = F)
+
+
+h_uav_top_res <- resample(h_uav_top, h_s2_top, method = "bilinear")
+
+h_uav_sd_res <- resample(h_uav_sd, h_s2_top, method = "bilinear")
+
+
+
+# crop to area of interest
+
+aoi <- shapefile("aoi.shp")
+
+h_uav_top_res <- mask(h_uav_top_res, aoi)
+
+h_uav_sd_res <- mask(h_uav_sd_res, aoi)
+
+h_s2_top <- mask(h_s2_top, aoi)
+
+h_s2_sd <- mask(h_s2_sd, aoi)
+
+
+h_uav_top_res <- crop(h_uav_top_res, aoi)
+
+h_uav_sd_res <- crop(h_uav_sd_res, aoi)
+
+h_s2_top <- crop(h_s2_top, aoi)
+
+h_s2_sd <- crop(h_s2_sd, aoi)
+
[…]

### 2023-07-20 submission_3/uav_s2_height_comparison.r

@@ -3,25 +3,25 @@ library(raster)
 # set your wd to where you have the downloaded data
 setwd()
 
-
+userpath <- "C:/Users/fabrikat/Documents/submission3/"
 
 # Load the two raster files
 
-h_uav_dsm <- raster("uav/cloudcompare_dsm_v2.tif")
+h_uav_dsm <- raster(paste(userpath, "uav/cloudcompare_dsm_v2.tif", sep=""))
 
 crs(h_uav_dsm) <- ("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs")
 
 
-h_uav_dsm <- raster("uav/DJI_202305311408_001_crane-2022-05_dem.tif")
+h_uav_dsm <- raster(paste(userpath, "/uav/DJI_202305311408_001_crane-2022-05_dem.tif", sep=""))
 
 # set 0 to NA
 
 h_uav_dsm <- reclassify(h_uav_dsm, cbind(0, NA))
 
 
-h_s2_top <- raster("sentinel_2_height/ETH_GlobalCanopyHeight_DE_CHM_clip.tif")
+h_s2_top <- raster(paste(userpath, "sentinel_2_height/ETH_GlobalCanopyHeight_DE_CHM_clip.tif", sep=""))
 
-h_s2_sd <- raster("sentinel_2_height/ETH_GlobalCanopyHeight_DE_STD_clip.tif")
+h_s2_sd <- raster(paste(userpath, "sentinel_2_height/ETH_GlobalCanopyHeight_DE_STD_clip.tif", sep=""))
 
 
 plot(h_uav_dsm-147)
@@ -49,7 +49,7 @@ h_uav_sd_res <- resample(h_uav_sd, h_s2_top, method = "bilinear")
 
 # crop to area of interest
 
-aoi <- shapefile("aoi.shp")
+aoi <- shapefile(paste(userpath, "shapefile/polygon.shp"))
 
 h_uav_top_res <- mask(h_uav_top_res, aoi)
 
@@ -68,32 +68,83 @@ h_s2_top <- crop(h_s2_top, aoi)
 
 h_s2_sd <- crop(h_s2_sd, aoi)
 
+h_res <- h_uav_top_res - h_s2_top
+
 # create plotting space
 par(mfrow = c(2, 2))
 
 # Plot 1
-plot(h_uav_top_res, main = "h_uav_top_res")
+plot(h_uav_top_res, main = "UAV height estimation")
 
[…]

### 2023-07-20 submission_4/submission_4.qmd
Adds first draft of submission 4

@@
 ---
-title: "submission_4"
+title: "Comparison of canopy height mesurment between TLS and Sentinel2 data"
 subtitle: "Documentation"
 subsubtitle: "Leaf traits measurements"
-author: | 
-    | **Group 4** 
-    | Denis Streitmatter 
-    | Marie-Louise Korte 
-    | Louis Trinkle 
-    | Maxwell Sivertsen 
-    | Fabian Götz 
-    | Lukas Erzfeld
+# Ok I reallly dont know why this does not work, 
+# author: | 
+#     | **Group 4** 
+#     | Denis Streitmatter 
+#     | Marie-Louise Korte 
+#     | Louis Trinkle 
+#     | Maxwell Sivertsen 
+#     | Fabian Götz 
+#     | Lukas Erzfeld
+author:  
+    # - **Group 4** 
+    - Denis Streitmatter 
+    - Marie-Louise Korte 
+    - Louis Trinkle 
+    - Maxwell Sivertsen 
+    - Fabian Götz 
+    - Lukas Erzfeld
 date: last-modified
@@
         theme: united
-        toc: true
+        toc: false
         toc-depth: 2
@@
 
-**Insert some description and the data of the plot. I wasn't there, so probably someone else should do it**
+**Study conditions:**
 
-## Algorithm idea
+- Sunny weather with a few clouds, ~ 23°
+- Location in the Leipzig Auwald, near the canopy crane (exact location can be seen in the figures)
+- Tools: terrestrial laser scanning (TLS), QGIS, Cloud Compare, R
+
+**Methodology**:
+
+1. TLS measurement was carried out at breast height
+2. Preprocessing of the TLS data via CloudCompare^[See [https://www.cloudcompare.org/](https://www.cloudcompare.org/)]
+3. QGIS was used for visualizing the sentinel data and creating the area of interest
+4. `R` was used to resample the TLS data to match the resolution of the Sentinel data
+5. `R` was also used to calculate the residuals and generally producing all figures in this document (importing the results of step 2, 3 and 4)
 
@@
 
-*First the general heigths:*
+The TLS height measurement is shown in @fig-tls-heights.
+It can be seen in the left plot that the TLS data is potentially far more fine grained than the data of the Senitnel2 satellite (see @fig-sentinel-height).
+For proper comparison the TLS got resampled to match the resolution of the sentiel data (see @fig-tls-heights)
+The difference in granularity can also been seen in the comparison of distribution of height and the standard derivations (see @fig-comparison-histograms): 
+While the height of TLS data is distributed evenly about the the whole height spectrum, the sentinel 2 data maps the whole area to a span of around 30 meters. 
+This can also be seen in @fig-sentinel-height, which uses the same colormap as @fig-tls-heights, but is very uniformly distributed.
 
-![Height calculation of TLS (resampled to resolution of the sentinel data)](TLS_plots/figures/chm_height.png)
+![Height calculation of TLS (left raw, right resampled to resolution of the sentinel data)](TLS_plots/figures/chm_tls_height_both_res.png){#fig-tls-heights}
 
-### Residual computation
+![Height of the Sentinel 2 data](TLS_plots/figures/sentinel_height.png){#fig-sentinel-height}
 
-In this section we compute the residuals of the
+![Height distribution and standard derivation of both Sentinel and TLS data as histograms](TLS_plots/figures/histograms.png){#fig-comparison-histograms}
+
+The reason for this is simple and twofold:
+
+1.
+**Resolution of satellite data**: The sheer distance of the satellite results in far lower resolution, resulting in "spacial aggregation", mapping a calculated value to a far larger area than the TLS data.
+2.
+**Way of measurement**: While the TLS scans from below and can therefor see the lower trees, Sentinel 2 measures from (very far) above, occluding the lower trees.
+This results in an overestimation of tree heights.
+But vice versa, the TLS data probably underestimates the canopy height, since lower trees can occlude the actual canopy.
+
+{{< pagebreak >}}
+
+### Comparison via residuals
+
+We computed the residuals by using the far more fine grained TLS data as truth and subtract the sentiel data from it. The result can be see in @fig-residuals-stats. 
+But there are a few things to consider:
+1. As briefly explained in the prevoius section, lower branches can occlude the actual canopy, resulting in an underestiation of the canopy height
+2. Also the further away from the device a tree is, the more likely it is to be (partly) occluded by closer objects.
+
+As expected, the Sentinel 2 data generally overestimates the canopy height, with a far larger part having negative residuals (colored red in @fig-residuals-stats). 
+Interestingly a small section in the map (bottom right, the area with the highest trees according to TLS data)
+
+![Residuals of the comparison.
+The far more precise TLS data serves here as the truth.
+Left is a map of residuals, right is a histogram of the distribution](TLS_plots/figures/residuals_stats.png){#fig-residuals-stats}
+
+## Conclusion
+
+In conclusion, it is shown that TLS and satellite data differ greatly in terms of granularity and variance, which is to no surprise. 
+Of course, the big advantage of using satellites, covering waste amout of area, is looked over in this approach to make it compareable to the TLS by selecting only such a small area.
+Furthermore, it is not entirely clear how good the values from the TLS actually are. 
+We used it as ground truth for calculating the residuals from the Sentinel data, but there are actually a few problems with this method, as explained before. 
+An interesting approach would be comparing TLS data to UAV data, which could show the possible underestimation of the TLS.
+
+So all in all, satellites are well suited for measuring vast areas, but lack the necessary precision to be used in pprecise, local measurements.
+On the other hand, TLS also has some systematic problems, which may be overcome by certain methods (multiple scanning from different locations and merge of the results, scanning at a higher level etc.).

### 2023-07-20 submission_4/TLS_plots/TLS_plots.R

@@ -14,10 +14,10 @@ library(rayshader)
 library(rgdal)
 library(sf)
 library(ggplot2)
+library(tidyverse)
 
-
-#setwd("/Users/marie-louisekorte/Documents/Uni Leipzig/Semester 2/Ground Truthing/analysis/TLS")
-setwd(getSrcDirectory(function(){})[1])
+setwd("/home/denis/workspace/uni/ground_truthing/submission_4/TLS_plots")
+# setwd(getSrcDirectory(function(){})[1])
 
 ### colorbar 
 col <- height.colors(25)
@@ -25,7 +25,6 @@ col <- height.colors(25)
 ### setting an extent to tls raster
 aoi <- shapefile("shapefile/polygon.shp")
 aoi
-
 
 ### TLS canopy height model
 chm_IEP_new <- raster('chm_IEP.tif')
@@ -72,6 +71,11 @@ hist(h_s2_top)
 
 plot(h_s2_sd)
 hist(h_s2_sd)
+
+chm_residuals_low_res <- h_tls_top_res - h_s2_top
+
+df_residuals_low_res <- as.data.frame(chm_residuals_low_res, xy = TRUE)
+colnames(df_residuals_low_res) <- c("x", "y", "value")
 
 
 ################################################################################################
@@ -152,6 +156,8 @@ ggplot(df_h_tls_sd_res, aes(x = x, y = y, fill = value)) +
        fill = "Standard Deviation")
 ggsave("figures/chm_sd.png", dpi = 300)
 
+
+
 ## sentinel
 df_h_s2_top <- as.data.frame(h_s2_top, xy = TRUE)
 colnames(df_h_s2_top) <- c("x", "y", "value")
@@ -192,10 +198,66 @@ ggsave("figures/sentinel_sd.png", dpi = 300)
 png(file="figures/histograms.png", width=900, height=750, units='px')
 
 par(mfrow=c(2,2))
-hist(h_tls_top_res, main = "TLS Mean Canopy Height", col = '#568fda', ylim = c(0,70), xlim = c(0,35), breaks = 20)
-hist(h_s2_top, main = "Sentinel Mean Canopy Height", col = '#568fda', ylim = c(0,70), xlim = c(0,35), breaks = 1)
-hist(h_tls_sd_res, main = "TLS SD of Canopy Height", col = "#038203", ylim = c(0,120),  xlim = c(0,12), breaks =20)
[…]

### 2023-07-20 submission_3/UAV+S2.qmd
Adds redered pdf of submission 3

@@
 subtitle: "Documentation"
-author: "Fabian Götz"
+author:  
+    # - **Group 4** 
+    - Denis Streitmatter 
+    - Marie-Louise Korte 
+    - Louis Trinkle 
+    - Maxwell Sivertsen 
+    - Fabian Götz 
+    - Lukas Erzfeld
 date: 6/24/23

### 2023-07-20 submission_3/UAV+S2.qmd
Adds links to src code

--- 2023-07-20 submission_3/UAV+S2.qmd
@@
 Specifically, one can see that Sentinel-2 does not resolve accurately enough to detect locations with low tree heights.
-Thus, the values of Sentinel-2 are too high on average.
+Thus, the values of Sentinel-2 are too high on average.^[The source code for all plots can be found at [https://github.com/yum-yab/ground_truthing_module/blob/main/submission_3/uav_s2_height_comparison.r](https://github.com/yum-yab/ground_truthing_module/blob/main/submission_3/uav_s2_height_comparison.r)]
 

### 2023-07-20 submission_4/submission_4.qmd
Adds links to src code

@@
 ---
-title: "Comparison of canopy height mesurment between TLS and Sentinel2 data"
+title: "Comparison of canopy height measurement between TLS and Sentinel2 data"
 subtitle: "Documentation"
@@
 4. `R` was used to resample the TLS data to match the resolution of the Sentinel data
-5. `R` was also used to calculate the residuals and generally producing all figures in this document (importing the results of step 2, 3 and 4)
+5.
+`R` was also used to calculate the residuals and generally producing all figures in this document (importing the results of step 2, 3 and 4)^[The code for everything can be found here [https://github.com/yum-yab/ground_truthing_module/blob/main/submission_4/TLS_plots/TLS_plots.R](https://github.com/yum-yab/ground_truthing_module/blob/main/submission_4/TLS_plots/TLS_plots.R)]
 

### 2023-07-27 submission_4/submission_4.qmd
Finishes missing sentence
smaller AOI

@@
 
-This section lists, describes and tries to explain the plots and graphs generated by the algorithm described in the previous section.
+This section lists, describes and explains the plots and graphs generated by the algorithm described in the previous section.
 
@@
 The TLS height measurement is shown in @fig-tls-heights.
-It can be seen in the left plot that the TLS data is potentially far more fine grained than the data of the Senitnel2 satellite (see @fig-sentinel-height).
-For proper comparison the TLS got resampled to match the resolution of the sentiel data (see @fig-tls-heights)
+It can be seen in the left plot that the TLS data is far more fine grained than the data of the Senitnel2 satellite (see @fig-sentinel-height).
+For better comparison the TLS got resampled to match the resolution of the sentiel data (see @fig-tls-heights)
 The difference in granularity can also been seen in the comparison of distribution of height and the standard derivations (see @fig-comparison-histograms): 
-While the height of TLS data is distributed evenly about the the whole height spectrum, the sentinel 2 data maps the whole area to a span of around 30 meters. 
+While the height of TLS data is distributed evenly across the the whole height spectrum, the sentinel 2 data maps the whole area to around 30 meters. 
 This can also be seen in @fig-sentinel-height, which uses the same colormap as @fig-tls-heights, but is very uniformly distributed.
@@
 
-We computed the residuals by using the far more fine grained TLS data as truth and subtract the sentiel data from it. The result can be see in @fig-residuals-stats. 
+We computed the residuals by using the far more finely resolved TLS data as truth and subtract the sentiel data from it. The result can be seen in @fig-residuals-stats. 
 But there are a few things to consider:
@@
 As expected, the Sentinel 2 data generally overestimates the canopy height, with a far larger part having negative residuals (colored red in @fig-residuals-stats). 
-Interestingly a small section in the map (bottom right, the area with the highest trees according to TLS data)
+Interestingly only a small section in the map (bottom right, the area with the highest trees according to TLS data) is underestimated.
 
@@
 
-In conclusion, it is shown that TLS and satellite data differ greatly in terms of granularity and variance, which is to no surprise. 
-Of course, the big advantage of using satellites, covering waste amout of area, is looked over in this approach to make it compareable to the TLS by selecting only such a small area.
+In conclusion, it is shown that TLS and satellite data differ greatly in terms of resolution and variance, which is no surprise. 
+Of course, the big advantage of using satellites, that can cover much larger areas, is over looked in this approach to make it compareable to the TLS by selecting only a small area.
 Furthermore, it is not entirely clear how good the values from the TLS actually are. 
@@
 
-So all in all, satellites are well suited for measuring vast areas, but lack the necessary precision to be used in pprecise, local measurements.
+So all in all, satellites are well suited for measuring vast areas, but lack the necessary precision to be used in precise, local measurements.
 On the other hand, TLS also has some systematic problems, which may be overcome by certain methods (multiple scanning from different locations and merge of the results, scanning at a higher level etc.).

### 2023-07-27 submission_4/TLS_plots/TLS_plots.R

@@ -10,20 +10,19 @@ library(forestr)
 library(lidR)
 library(rayshader)
 library(RCSF)
-library(rayshader)
 library(rgdal)
 library(sf)
 library(ggplot2)
 library(tidyverse)
 
-setwd("/home/denis/workspace/uni/ground_truthing/submission_4/TLS_plots")
+setwd("D:/Master/2_sem/ground_trouthing/ground_truthing_module/submission_4/TLS_plots")
 # setwd(getSrcDirectory(function(){})[1])
 
 ### colorbar 
 col <- height.colors(25)
 
 ### setting an extent to tls raster
-aoi <- shapefile("shapefile/polygon.shp")
+aoi <- shapefile("shapefile/polygon_small.shp")
 aoi
 
 ### TLS canopy height model
